We want to determine whether viruses are more spread than the other classes

```{r libraries}
library(data.table)
library(cluster)
library(ggplot2)
library(corrplot)
library(plyr)
library(dplyr)
library(reshape2)
setwd("~/Documents/metaviro/experiments/140528_full_ncbi_resampling")
kmers=fread("full_ncbi_contigs_4k_v2_3mers.csv")
kmers_columns=colnames(kmers)[5:ncol(kmers)]
sequence_attributes=limma::strsplit2(kmers$sequence_description,"_")
kmers$species=factor(sequence_attributes[,1])
kmers$class=factor(sequence_attributes[,2])
kmers$contig_index=as.numeric(sequence_attributes[,4])
kmers$is_RC=sequence_attributes[,7]=="rev"
normalized_kmers=kmers[,kmers_columns,with=F]/kmers$sequence_length
```

* We downsample 

```{r }
downsample_idx=sample(1:nrow(normalized_kmers),2500)

kmers_ds=kmers[downsample_idx,]
kmers_ds$idx=1:nrow(kmers_ds)
normalized_kmers_ds=normalized_kmers[downsample_idx,]
dim(kmers_ds)
```

* We compute the pairwise distance matrix 

```{r }
pairwise_dist = dist(normalized_kmers_ds)
pairwise_dist_m=as.matrix(pairwise_dist)
pairwise_dist_m[lower.tri(pairwise_dist_m,diag=T)]<-NA

rownames(pairwise_dist_m)=kmers_ds$species
colnames(pairwise_dist_m)=kmers_ds$species
dim(pairwise_dist_m)
pairwise_dist_m=data.table(melt(pairwise_dist_m))
pairwise_dist_m=na.omit(pairwise_dist_m)


pairwise_dist_m$Var1=factor(pairwise_dist_m$Var1)
pairwise_dist_m$Var2=factor(pairwise_dist_m$Var2)

str(pairwise_dist_m)
# Add the class annotation 
class_annot=unique(kmers_ds[,list(species,class)])
setkey(class_annot,"species")


pairwise_dist_m$src_class=class_annot[J(pairwise_dist_m[,Var1])][,class]
pairwise_dist_m$tgt_class=class_annot[J(pairwise_dist_m[,Var2])][,class]

str(pairwise_dist_m)


```


* Averages

```{r }
avg_dist=pairwise_dist_m[,median(value),by=list(src_class,tgt_class)][order(V1)]
avg_dist=dcast(avg_dist,src_class~tgt_class)
rownames(avg_dist) = avg_dist$src_class
avg_dist$src_class<-NULL
corrplot(as.matrix(avg_dist)*100,is.corr=F)
heatmap(as.numeric(avg_dist),is.corr=F)

```
No information, except that euk and bact are far apart, and that avg euk - euk distance is the largest of the diags 

* pair wise boxplots 

```{r }
ggplot(pairwise_dist_m,aes(x=src_class,y=value))+geom_boxplot()+facet_wrap(~tgt_class)

```

* Les copains autour (k-NN)

```{r }
pairwise_dist_m[,ranked_dist:=rank(value),by=Var1]

matched_friends=pairwise_dist_m[ranked_dist<=7]

```

We determine the most frequent class in the neighborhood 

```{r }


matched_friends[,vir_counts:={foo=table(tgt_class);foo["viruses"]},by=list(Var1,src_class)]
matched_friends[,bact_counts:={foo=table(tgt_class);foo["bact"]},by=list(Var1,src_class)]
matched_friends[,arch_counts:={foo=table(tgt_class);foo["archea"]},by=list(Var1,src_class)]
matched_friends[,euk_counts:={foo=table(tgt_class);foo["euk"]},by=list(Var1,src_class)]
matched_friends[,k_disagreeing:=vir_counts+ bact_counts+ arch_counts+ euk_counts]

matched_friends=unique(matched_friends[,list(Var1,src_class,vir_counts, bact_counts, arch_counts, euk_counts,k_disagreeing)])


matched_friends[src_class=="viruses",k_disagreeing:=k_disagreeing-vir_counts]
matched_friends[src_class=="archea",k_disagreeing:=k_disagreeing-arch_counts]
matched_friends[src_class=="bact",k_disagreeing:=k_disagreeing-bact_counts]
matched_friends[src_class=="euk",k_disagreeing:=k_disagreeing-euk_counts]


# matched_friends[,mean(k_disagreeing),by=src_class]

matched_friends[k_disagreeing>=5,.N,by=src_class]

ggplot(matched_friends,aes(x=src_class,fill=src_class,y=k_disagreeing))+geom_histogram()
ggplot(matched_friends,aes(fill=src_class,x=factor(k_disagreeing)))+geom_histogram()

matched_friends[k_disagreeing==4,.N,by=src_class]
matched_friends[,.N,by=src_class]

sum(matched_friends[,.N,by=list(src_class,k_disagreeing)][order(src_class,k_disagreeing)]$N)
```


```{r test}
ddply(matched_friends,.variables=c("Var1","src_class"),.fun=table)

find_most_frequent=function(vec){
	print(vec)
	sorted_table=sort(table(vec),decreasing=T)
	print(sorted_table)
	return(sorted_table[1])
}

matched_friends[1:10,find_most_frequent(tgt_class),by=list(Var1,src_class)]

confusionMatrix(matched_friends[,src_class],matched_friends[,V1])
```

* MCA analysis 

```{r }
library(FactoMineR)



res.mca=PCA(kmers_ds[,c("class",kmers_columns),with=F],quali.sup=c(1))
plot.PCA(res.mca,invisible=c("quali"),habillage=c(1),label="none")
plot.PCA(res.mca,invisible=c("quali"),habillage=c(1),axes=c(1,3),label="none")
plot.PCA(res.mca,invisible=c("quali"),habillage=c(1),axes=c(2,3),label="none")
dimdesc(res.mca)

```

We do a hierachical clustering 

```{r }
h_clust=HCPC(res.mca) #very slow due to drawing! 

clustered_data=data.table(h_clust$data.clust)
dcast(clustered_data[,.N,by=list(clust,class)],class~clust)

```
HCPC show no good association with the class labels 

* Other approaches, where we compute the distances to the class centroids (possibly having multiple class centroids)
