```{r }
library(data.table)
library(infotheo)
library(dplyr)
library(gridExtra)
library(reshape2)
library(ggplot2)
library(caret)

```

We load a large pattern 

```{r }
setwd("~/Documents/metaviro/experiments/140613_more_patterns/")
#spaced_kmers=fread("spaced_kmers/long_kmers_m300_n4.fa_101011010111.csv")

# spaced_kmers=fread("spaced_kmers/long_kmers_m300_n4.fa_111101000010101.csv")
spaced_kmers=fread("spaced_kmers/long_kmers_m300_n4.fa_11101011.csv")


seq_attributes=limma::strsplit2(spaced_kmers$sequence_description,"_")
spaced_kmers$class=factor(seq_attributes[,2])
spaced_kmers$species=factor(seq_attributes[,1])
spaced_kmers$pattern=factor(spaced_kmers$pattern)
spaced_kmers$kmer=factor(spaced_kmers$kmer)
```

What are the sparse k-mers ?

```{r }
kmer_by_species=spaced_kmers[,sum(count),by=list(kmer,class,species)]
# We binarize 
kmer_by_species$V1=1

# This give the number of unique species in which we find a kmer 
kmer_by_unique_species=kmer_by_species[,sum(V1),by=list(kmer,class)][order(V1)]

kmer_by_unique_species_m=dcast.data.table(kmer_by_unique_species,kmer~class,fill=0)

# We normalize by the number of species in each classes 
n_classes=spaced_kmers[,.N,by=list(class,species)][,.N,by=class]
n_classes_l=as.list(n_classes$N)
names(n_classes_l)=as.character(n_classes$class)

kmer_by_unique_species_m$bact = kmer_by_unique_species_m$bact/n_classes_l$bact 
kmer_by_unique_species_m$euk = kmer_by_unique_species_m$euk/n_classes_l$euk 
kmer_by_unique_species_m$archea = kmer_by_unique_species_m$archea/n_classes_l$archea 
kmer_by_unique_species_m$viruses = kmer_by_unique_species_m$viruses/n_classes_l$viruses 

kmer_by_unique_species_m[,sd:=sd(c(archea,bact,euk,viruses)),by=kmer]
kmer_by_unique_species_m[order(sd)]
g1=ggplot(kmer_by_unique_species_m,aes(x=bact,y=sd,colour=viruses))+geom_point()+geom_smooth()
g2=ggplot(kmer_by_unique_species_m,aes(x=euk,y=sd,colour=viruses))+geom_point()+geom_smooth()
g3=ggplot(kmer_by_unique_species_m,aes(x=archea,y=sd,colour=viruses))+geom_point()+geom_smooth()
g4=ggplot(kmer_by_unique_species_m,aes(x=viruses,y=sd,colour=euk))+geom_point()+geom_smooth()
grid.arrange(g1,g2,g3,g4)

ggplot(kmer_by_unique_species_m,aes(x=archea,y=bact,colour=(viruses>=0.4)))+geom_point()+geom_smooth()
```


# Feature selection 

cf 
	http://www.hpl.hp.com/techreports/2007/HPL-2007-16R1.pdf
	http://nlp.stanford.edu/IR-book/html/htmledition/assessing-as-a-feature-selection-methodassessing-chi-square-as-a-feature-selection-method-1.html
	http://blog.datumbox.com/using-feature-selection-methods-in-text-classification/
	http://courses.ischool.berkeley.edu/i256/f06/papers/yang97comparative.pdf


Number of pos / neg docs for each feature at the species level 

```{r }
kmer_by_unique_species_count=dcast.data.table(kmer_by_unique_species,kmer~class,fill=0)
ggplot(kmer_by_unique_species_count,aes(x=archea+bact+euk,y=viruses))+geom_point()
```

We compute the chi^2 stat 

```{r }
kmer_viral_contingency=kmer_by_unique_species_count[,list(
	n_viral_no=n_classes_l$viruses - viruses, # N viral species without the k-mer 
	n_viral=viruses, # N viral species with the k-mer
	n_no_viral=archea+bact+euk, # N non-viral with the k-mer 
	n_no_viral_no=sum(unlist(n_classes_l))-n_classes_l$viruses - (archea+bact+euk) # N non-viral without the k-mer 
	),by=kmer]

kmer_viral_chi2=kmer_viral_contingency[,list(viral_chi2=chisq.test(matrix(c(n_viral_no,n_viral,n_no_viral_no,n_no_viral),byrow=T,ncol=2))$statistic),by=kmer]


kmer_euk_contingency=kmer_by_unique_species_count[,list(
  n_euk_no=n_classes_l$euk - euk, # N euk species without the k-mer 
	n_euk=euk, # N euk species with the k-mer
	n_no_euk=archea+bact+viruses, # N non-euk with the k-mer 
	n_no_euk_no=sum(unlist(n_classes_l))-n_classes_l$euk - (archea+bact+viruses) # N non-euk without the k-mer 
	),by=kmer]

kmer_euk_chi2=kmer_euk_contingency[,list(euk_chi2=chisq.test(matrix(c(n_euk_no,n_euk,n_no_euk_no,n_no_euk),byrow=T,ncol=2))$statistic),by=kmer]


kmer_bact_contingency=kmer_by_unique_species_count[,list(
  n_bact_no=n_classes_l$bact - bact, # N bact species without the k-mer 
	n_bact=bact, # N bact species with the k-mer
	n_no_bact=archea+bact+viruses, # N non-bact with the k-mer 
	n_no_bact_no=sum(unlist(n_classes_l))-n_classes_l$bact - (archea+euk+viruses) # N non-bact without the k-mer 
	),by=kmer]

kmer_bact_chi2=kmer_bact_contingency[,list(bact_chi2=chisq.test(matrix(c(n_bact_no,n_bact,n_no_bact_no,n_no_bact),byrow=T,ncol=2))$statistic),by=kmer]


kmer_archea_contingency=kmer_by_unique_species_count[,list(
  n_archea_no=n_classes_l$archea - archea, # N archea species without the k-mer 
	n_archea=archea, # N archea species with the k-mer
	n_no_archea=archea+archea+viruses, # N non-archea with the k-mer 
	n_no_archea_no=sum(unlist(n_classes_l))-n_classes_l$archea - (archea+euk+viruses) # N non-archea without the k-mer 
	),by=kmer]

kmer_archea_chi2=kmer_archea_contingency[,list(archea_chi2=chisq.test(matrix(c(n_archea_no,n_archea,n_no_archea_no,n_no_archea),byrow=T,ncol=2))$statistic),by=kmer]



kmer_by_unique_species_count=merge(merge(merge(merge(kmer_by_unique_species_count,kmer_viral_chi2,by="kmer"),kmer_euk_chi2,by="kmer"),kmer_bact_chi2,by='kmer'),kmer_archea_chi2,by="kmer")

g1=ggplot(kmer_by_unique_species_count,aes(x=archea+bact+euk,y=viruses,colour=viral_chi2))+geom_point()
g2=ggplot(kmer_by_unique_species_count,aes(x=archea+bact+viruses,y=euk,colour=euk_chi2))+geom_point()
g3=ggplot(kmer_by_unique_species_count,aes(x=archea+euk+viruses,y=bact,colour=bact_chi2))+geom_point()
g4=ggplot(kmer_by_unique_species_count,aes(x=euk+bact+viruses,y=archea,colour=archea_chi2))+geom_point()
grid.arrange(g1,g2,g3,g4)
```

Instead of a one-vs-all chi2, we compute the chi2 with all classes assignment 

```{r }

kmer_all_contingency=kmer_by_unique_species_count[,list(
  	n_archea_no=n_classes_l$archea - archea, # N archea species without the k-mer 
  	n_bact_no=n_classes_l$bact - bact, # N archea species without the k-mer 
  	n_euk_no=n_classes_l$euk - euk, # N archea species without the k-mer 
  	n_viruses_no=n_classes_l$viruses - viruses, # N archea species without the k-mer 
  	n_archea=archea, # N archea species without the k-mer 
  	n_bact=bact, # N archea species without the k-mer 
  	n_euk=euk, # N archea species without the k-mer 
  	n_viruses=viruses # N archea species without the k-mer 
	),by=kmer]

kmer_chi2=kmer_all_contingency[,list(global_chi2=chisq.test(matrix(c(n_archea_no, n_bact_no, n_euk_no, n_viruses_no, n_archea, n_bact, n_euk, n_viruses),byrow=T,ncol=4))$statistic),by=kmer]
#setnames(kmer_chi2,c("kmer","global_chi2"))
kmer_by_unique_species_count=merge(kmer_by_unique_species_count,kmer_chi2,by='kmer')

#10.83 == signif at the 0.001 level a
g1=ggplot(kmer_by_unique_species_count,aes(x=archea+bact+euk,y=viruses,colour=interaction(global_chi2>=10.83,viral_chi2>=10.83)))+geom_point()+geom_abline(slope=1)
g2=ggplot(kmer_by_unique_species_count,aes(x=archea+bact+viruses,y=euk,colour=interaction(global_chi2>=10.83,euk_chi2>=10.83)))+geom_point()+geom_abline(slope=1)
g3=ggplot(kmer_by_unique_species_count,aes(x=archea+euk+viruses,y=bact,colour=interaction(global_chi2>=10.83,bact_chi2>=10.83)))+geom_point()+geom_abline(slope=1)
g4=ggplot(kmer_by_unique_species_count,aes(x=euk+bact+viruses,y=archea,colour=interaction(global_chi2>=10.83,archea_chi2>=10.83)))+geom_point()+geom_abline(slope=1)
grid.arrange(g1,g2,g3,g4)

```


We compare classifiers using 64 random features vs classifier using top 64 chi2 features 


```{r}
random_features=sample(levels(spaced_kmers$kmer),size=64)
top_features=as.character(tail(kmer_by_unique_species_count[order(global_chi2)],n=64)$kmer)


random_feat_matrix=dcast.data.table(spaced_kmers[kmer %in% random_features],sequence_description+class~kmer,value.var="count",fill=0)
top_feat_matrix=dcast.data.table(spaced_kmers[kmer %in% top_features],sequence_description+class~kmer,value.var="count",fill=0)
count_cols_rand=colnames(random_feat_matrix)[-(1:2)]
count_cols_top=colnames(top_feat_matrix)[-(1:2)]

# We make a balanced data set 

sample_row = function(df,size){df[sample(1:nrow(df),size=size),]}
balanced_size=min(table(top_feat_matrix$class))
top_feat_matrix_balanced=rbind(
    top_feat_matrix[class=="archea"] %.% sample_row(balanced_size),
    top_feat_matrix[class=="bact"] %.% sample_row(balanced_size),
    top_feat_matrix[class=="euk"] %.% sample_row(balanced_size),
    top_feat_matrix[class=="viruses"] %.% sample_row(balanced_size))

```

```{r}
library(caret)

set.seed(3456)
trainIndex <- createDataPartition(top_feat_matrix$class, p = .8, list = FALSE, times = 1)

seqTrain <- top_feat_matrix[ trainIndex[,1]][,c("class",count_cols_top),with=F]
seqTest  <- top_feat_matrix[-trainIndex[,1],c("class",count_cols_top),with=F]

fitControl <- trainControl(    method = "repeatedcv", number = 3, repeats = 1)

svmGrid=expand.grid(sigma=10^(-5:-1),C=c(1,2,3))

svmTrainFit <- train(class ~ ., data = seqTrain,
                 method = "svmRadial",
                 trControl = fitControl,
                 tuneGrid=svmGrid,
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = T)


predictions=predict(svmTrainFit,newdata=seqTest[,count_cols_top,with=F])
confusionMatrix(predictions,seqTest$class)

```
We compare against a balanced dataset 

```{r }
set.seed(3456)
trainIndex <- createDataPartition(top_feat_matrix_balanced$class, p = .8, list = FALSE, times = 1)

seqTrain <- top_feat_matrix_balanced[ trainIndex[,1]][,c("class",count_cols_top),with=F]
seqTest  <- top_feat_matrix_balanced[-trainIndex[,1],c("class",count_cols_top),with=F]

fitControl <- trainControl(    method = "repeatedcv", number = 3, repeats = 1)

svmGrid=expand.grid(sigma=10^(-2:-1),C=c(0.1,1,2,3))

svmTrainFit <- train(class ~ ., data = seqTrain,
                 method = "svmRadial",
                 trControl = fitControl,
                 tuneGrid=svmGrid,
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = T)


predictions=predict(svmTrainFit,newdata=seqTest[,count_cols_top,with=F])
confusionMatrix(predictions,seqTest$class)


```

We compare against SVM gamma 90 3-mers

```{r}
all_3_mers=fread('spaced_kmers/long_kmers_m300_n4.fa_111.csv')


seq_attributes=limma::strsplit2(all_3_mers$sequence_description,"_")
all_3_mers$class=factor(seq_attributes[,2])
all_3_mers$species=factor(seq_attributes[,1])
all_3_mers$pattern=factor(all_3_mers$pattern)
all_3_mers$kmer=factor(all_3_mers$kmer)

all_3_mers_counts=dcast.data.table(all_3_mers,class+sequence_description~kmer,value.var='count',fill=0)
count_cols_3mers=colnames(all_3_mers_counts)[-(1:2)]

set.seed(3456)
trainIndex <- createDataPartition(all_3_mers_counts$class, p = .8, list = FALSE, times = 1)

seqTrain <- all_3_mers_counts[ trainIndex[,1]][,c("class",count_cols_3mers),with=F]
seqTest  <- all_3_mers_counts[-trainIndex[,1],c("class",count_cols_3mers),with=F]

fitControl <- trainControl(    method = "repeatedcv", number = 3, repeats = 2)

svmGrid=expand.grid(sigma=10^(-2:-1),C=c(0.1,1,2,3))

svmTrainFit3Mers <- train(class ~ ., data = seqTrain,
                 method = "svmRadial",
                 trControl = fitControl,
                 tuneGrid=svmGrid,
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = T)

predictions=predict(svmTrainFit3Mers,newdata=seqTrain[,count_cols_3mers,with=F])
confusionMatrix(predictions,seqTrain$.outcome)

predictions=predict(svmTrainFit3Mers,newdata=seqTest[,count_cols_3mers,with=F])
confusionMatrix(predictions,seqTest$class)

```

SVM over 4-mers 

```{r }

all_4_mers=fread('spaced_kmers/long_kmers_m300_n4.fa_1111.csv')


seq_attributes=limma::strsplit2(all_4_mers$sequence_description,"_")
all_4_mers$class=factor(seq_attributes[,2])
all_4_mers$species=factor(seq_attributes[,1])
all_4_mers$pattern=factor(all_4_mers$pattern)
all_4_mers$kmer=factor(all_4_mers$kmer)

all_4_mers_counts=dcast.data.table(all_4_mers,class+sequence_description~kmer,value.var='count',fill=0)
count_cols_4mers=colnames(all_4_mers_counts)[-(1:2)]

set.seed(3456)
trainIndex <- createDataPartition(all_4_mers_counts$class, p = .8, list = FALSE, times = 1)

seqTrain <- all_4_mers_counts[ trainIndex[,1]][,c("class",count_cols_4mers),with=F]
seqTest  <- all_4_mers_counts[-trainIndex[,1],c("class",count_cols_4mers),with=F]

fitControl <- trainControl(    method = "repeatedcv", number = 3, repeats = 2)

svmGrid=expand.grid(sigma=10^(-2:-1),C=c(0.1,1,2,3))

svmTrainFit4Mers <- train(class ~ ., data = seqTrain,
                 method = "svmRadial",
                 trControl = fitControl,
                 tuneGrid=svmGrid,
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = T)

predictions=predict(svmTrainFit4Mers,newdata=seqTrain[,count_cols_4mers,with=F])
confusionMatrix(predictions,seqTrain$.outcome)

predictions=predict(svmTrainFit4Mers,newdata=seqTest[,count_cols_4mers,with=F])
confusionMatrix(predictions,seqTest$class)
```

We retrain on a balanced version of the 4-mers dataset 

```{r }
balanced_size=min(table(all_4_mers_counts$class))
all_4_mers_counts_balanced=rbind(
    all_4_mers_counts[class=="archea"] %.% sample_row(balanced_size),
    all_4_mers_counts[class=="bact"] %.% sample_row(balanced_size),
    all_4_mers_counts[class=="euk"] %.% sample_row(balanced_size),
    all_4_mers_counts[class=="viruses"] %.% sample_row(balanced_size))



set.seed(3456)
trainIndex <- createDataPartition(all_4_mers_counts_balanced$class, p = .8, list = FALSE, times = 1)

seqTrain <- all_4_mers_counts_balanced[ trainIndex[,1]][,c("class",count_cols_4mers),with=F]
seqTest  <- all_4_mers_counts_balanced[-trainIndex[,1],c("class",count_cols_4mers),with=F]

svmTrainFit4Mers_balanced <- train(class ~ ., data = seqTrain,
                 method = "svmRadial",
                 trControl = trainControl(method = "none"),
                 tuneGrid=data.frame(sigma=0.01,C=2),
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = T)

predictions=predict(svmTrainFit4Mers_balanced,newdata=seqTest[,count_cols_4mers,with=F])
confusionMatrix(predictions,seqTest$class)

```
This 4-mers SVM over balanced data yields the best accuracy seen so far (0.90, Kappa=0.90), almost perfect classification



We compare against a round-robin schema where each class can suggest features 

```{r}

top_features = unique(c(
  as.character(head(kmer_by_unique_species_count[order(viral_chi2,decreasing=T),kmer],n=16)),
  as.character(head(kmer_by_unique_species_count[order(bact_chi2,decreasing=T),kmer],n=16)),
  as.character(head(kmer_by_unique_species_count[order(euk_chi2,decreasing=T),kmer],n=16)),
  as.character(head(kmer_by_unique_species_count[order(archea_chi2,decreasing=T),kmer],n=16))))
  
  
top_feat_matrix=dcast.data.table(spaced_kmers[kmer %in% top_features],sequence_description+class~kmer,value.var="count",fill=0)
count_cols_top=colnames(top_feat_matrix)[-(1:2)]

set.seed(3456)
trainIndex <- createDataPartition(top_feat_matrix$class, p = .8, list = FALSE, times = 1)

seqTrain <- top_feat_matrix[ trainIndex[,1]][,c("class",count_cols_top),with=F]
seqTest  <- top_feat_matrix[-trainIndex[,1],c("class",count_cols_top),with=F]

fitControl <- trainControl(    method = "repeatedcv", number = 3, repeats = 2)

svmGrid=expand.grid(sigma=c(1,10,90),C=c(1,10))

svmTrainFitTop <- train(class ~ ., data = seqTrain,
                 method = "svmRadial",
                 trControl = fitControl,
                 #tuneGrid=svmGrid,
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = T)

predictions=predict(svmTrainFitTop,newdata=seqTest[,count_cols_top,with=F])
confusionMatrix(predictions,seqTest$class)

```
Not better.

We perform a scale centering on the 3-mers 

```{r}
set.seed(3456)
trainIndex <- createDataPartition(all_3_mers_counts$class, p = .8, list = FALSE, times = 1)

seqTrain <- all_3_mers_counts[ trainIndex[,1]][,c("class",count_cols_3mers),with=F]
seqTest  <- all_3_mers_counts[-trainIndex[,1],c("class",count_cols_3mers),with=F]

preProcValues <- preProcess(seqTrain[,count_cols_3mers,with=F], method = c("center", "scale"))
trainTransformed <- predict(preProcValues, seqTrain[,count_cols_3mers,with=F])
trainTransformed$class = seqTrain$class
testTransformed <- predict(preProcValues, seqTest[,count_cols_3mers,with=F])
testTransformed$class = seqTest$class



svmTrainFit3Mers <- train(class ~ ., data = trainTransformed,
                 method = "svmRadial",
                 trControl = fitControl,
                 #tuneGrid=svmGrid,
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = T)


predictions=predict(svmTrainFit3Mers,newdata=testTransformed[,count_cols_3mers])
confusionMatrix(predictions,testTransformed$class)

```

# We try on a more complete data sets (more viral species, more non viral sequences)

```{r }
complete_dset_4mers=fread("spaced_kmers/long_kmers_mVar_n30.fa_1111.csv")
seq_attributes=limma::strsplit2(complete_dset_4mers$sequence_description,"_")
complete_dset_4mers$class=factor(seq_attributes[,2])
complete_dset_4mers$species=factor(seq_attributes[,1])
complete_dset_4mers$pattern=factor(complete_dset_4mers$pattern)
complete_dset_4mers$kmer=factor(complete_dset_4mers$kmer)

complete_dset_4mers_counts=dcast.data.table(complete_dset_4mers,class+sequence_description~kmer,value.var='count',fill=0)
count_cols_4mers=colnames(complete_dset_4mers_counts)[-(1:2)]
table(complete_dset_4mers_counts$class)

# We balance in terms of sequences 
balanced_size=min(table(complete_dset_4mers_counts$class))
complete_dset_4mers_counts_bal=rbind(
    complete_dset_4mers_counts[class=="archea"] %.% sample_row(balanced_size),
    complete_dset_4mers_counts[class=="bact"] %.% sample_row(balanced_size),
    complete_dset_4mers_counts[class=="euk"] %.% sample_row(balanced_size),
    complete_dset_4mers_counts[class=="viruses"] %.% sample_row(balanced_size))


set.seed(3456)
trainIndex <- createDataPartition(complete_dset_4mers_counts_bal$class, p = .6, list = FALSE, times = 1)

seqTrain <- complete_dset_4mers_counts_bal[ trainIndex[,1]][,c("class",count_cols_4mers),with=F]
seqTest  <- complete_dset_4mers_counts_bal[-trainIndex[,1],c("class",count_cols_4mers),with=F]


svmTrainFit4Mers_balanced_complete <- train(class ~ ., data = seqTrain,
                 method = "svmRadial",
                 trControl = trainControl(method = "none"),
                 tuneGrid=data.frame(sigma=0.01,C=2),
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = T)

predictions=predict(svmTrainFit4Mers_balanced_complete,newdata=seqTest[,count_cols_4mers,with=F])
confusionMatrix(predictions,seqTest$class)

```

For this dataset we determine whether normalization helps 

```{r }
preProcValues <- preProcess(seqTrain[,count_cols_4mers,with=F], method = c("center", "scale"))
trainTransformed <- as.data.frame(predict(preProcValues, seqTrain[,count_cols_4mers,with=F]))
trainTransformed$class = seqTrain$.outcome
testTransformed <- predict(preProcValues, seqTest[,count_cols_4mers,with=F])
testTransformed$class = seqTest$class

svmTrainFit4Mers_balanced_complete_norm <- train(class ~ ., data = trainTransformed,
                 method = "svmRadial",
                 trControl = trainControl(method = "none"),
                 tuneGrid=data.frame(sigma=0.01,C=2),
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = T)

predictions=predict(svmTrainFit4Mers_balanced_complete_norm,newdata=testTransformed[,count_cols_4mers])
confusionMatrix(predictions,seqTest$class)
```
Nope.

We determine whether length normalization helps 

```{r }
seqTrain_length=apply(seqTrain[,count_cols_4mers,with=F],1,sum)
seqTrain_length_normalized=seqTrain[,count_cols_4mers,with=F]/seqTrain_length
seqTrain_length_normalized$class=seqTrain$.outcome

seqTest_length=apply(seqTest[,count_cols_4mers,with=F],1,sum)
seqTest_length_normalized=seqTest[,count_cols_4mers,with=F]/seqTest_length
seqTest_length_normalized$class=seqTest$class

svmTrainFit4Mers_balanced_complete_lnorm <- train(class ~ ., data = seqTrain_length_normalized,
                 method = "svmRadial",
                 trControl = trainControl(method = "none"),
                 tuneGrid=data.frame(sigma=0.01,C=2),
                 verbose = T)

predictions=predict(svmTrainFit4Mers_balanced_complete_lnorm,newdata=seqTest_length_normalized)
confusionMatrix(predictions,seqTest_length_normalized$class)
```

Yes it does.

We compare the SVM 4-mers with an SVM built on chi2 signif features over non-sparse spaced k-mers

```{r }
spaced_kmers=fread("spaced_kmers/long_kmers_mVar_n30.fa_11101.csv")

seq_attributes=limma::strsplit2(spaced_kmers$sequence_description,"_")
spaced_kmers$class=factor(seq_attributes[,2])
spaced_kmers$species=factor(seq_attributes[,1])
spaced_kmers$pattern=factor(spaced_kmers$pattern)
spaced_kmers$kmer=factor(spaced_kmers$kmer)
rm(seq_attributes)



complete_dset_4_spaced_mers_counts=dcast.data.table(spaced_kmers,class+sequence_description~kmer,value.var='count',fill=0)
count_cols_4mers=colnames(complete_dset_4_spaced_mers_counts)[-(1:2)]
table(complete_dset_4_spaced_mers_counts$class)

# We balance in terms of sequences 
balanced_size=min(table(complete_dset_4_spaced_mers_counts$class))
complete_dset_4_spaced_mers_counts_bal=rbind(
    complete_dset_4_spaced_mers_counts[class=="archea"] %.% sample_row(balanced_size),
    complete_dset_4_spaced_mers_counts[class=="bact"] %.% sample_row(balanced_size),
    complete_dset_4_spaced_mers_counts[class=="euk"] %.% sample_row(balanced_size),
    complete_dset_4_spaced_mers_counts[class=="viruses"] %.% sample_row(balanced_size))


set.seed(3456)
trainIndex <- createDataPartition(complete_dset_4_spaced_mers_counts_bal$class, p = .6, list = FALSE, times = 1)

seqTrain <- complete_dset_4_spaced_mers_counts_bal[ trainIndex[,1]][,c("class",count_cols_4mers),with=F]
seqTest  <- complete_dset_4_spaced_mers_counts_bal[-trainIndex[,1],c("class",count_cols_4mers),with=F]

seqTrain_length=apply(seqTrain[,count_cols_4mers,with=F],1,sum)
seqTrain_length_normalized=seqTrain[,count_cols_4mers,with=F]/seqTrain_length
seqTrain_length_normalized$class=seqTrain$class

seqTest_length=apply(seqTest[,count_cols_4mers,with=F],1,sum)
seqTest_length_normalized=seqTest[,count_cols_4mers,with=F]/seqTest_length
seqTest_length_normalized$class=seqTest$class




svmTrainFit4_spaced_Mers_balanced_complete <- train(class ~ ., data = seqTrain_length_normalized,
                 method = "svmRadial",
                 trControl = trainControl(method = "none"),
                 tuneGrid=data.frame(sigma=0.01,C=2),
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = T)

predictions=predict(svmTrainFit4_spaced_Mers_balanced_complete,newdata=seqTest_length_normalized[,count_cols_4mers,with=F])
confusionMatrix(predictions,seqTest$class)


```

We get 

  Confusion Matrix and Statistics

            Reference
  Prediction archea bact  euk viruses
     archea    1100   91   83     122
     bact        66 1082   42     135
     euk        141   95 1113     176
     viruses     77  116  146     951

  Overall Statistics
                                            
                 Accuracy : 0.767           
                   95% CI : (0.7556, 0.7781)
      No Information Rate : 0.25            
      P-Value [Acc > NIR] : < 2.2e-16       
                                            
                    Kappa : 0.6893          
   Mcnemar's Test P-Value : 7.686e-10       

  Statistics by Class:

                       Class: archea Class: bact Class: euk Class: viruses
  Sensitivity                 0.7948      0.7818     0.8042         0.6871
  Specificity                 0.9287      0.9415     0.9008         0.9184
  Pos Pred Value              0.7880      0.8166     0.7298         0.7372
  Neg Pred Value              0.9314      0.9283     0.9324         0.8980
  Prevalence                  0.2500      0.2500     0.2500         0.2500
  Detection Rate              0.1987      0.1954     0.2010         0.1718
  Detection Prevalence        0.2522      0.2393     0.2755         0.2330
  Balanced Accuracy           0.8618      0.8616     0.8525         0.8027




# 3-mers on the same data set 

```{r }

all_3mers=fread("spaced_kmers/long_kmers_mVar_n30.fa_111.csv")

seq_attributes=limma::strsplit2(all_3mers$sequence_description,"_")
all_3mers$class=factor(seq_attributes[,2])
all_3mers$species=factor(seq_attributes[,1])
all_3mers$pattern=factor(all_3mers$pattern)
all_3mers$kmer=factor(all_3mers$kmer)
rm(seq_attributes)



complete_dset_3mers_counts=dcast.data.table(all_3mers,class+sequence_description~kmer,value.var='count',fill=0)
count_cols_3mers=colnames(complete_dset_3mers_counts)[-(1:2)]
table(complete_dset_3mers_counts$class)

# We balance in terms of sequences 
balanced_size=min(table(complete_dset_3mers_counts$class))

set.seed(3456)
complete_dset_3mers_counts_bal=rbind(
    complete_dset_3mers_counts[class=="archea"] %.% sample_row(balanced_size),
    complete_dset_3mers_counts[class=="bact"] %.% sample_row(balanced_size),
    complete_dset_3mers_counts[class=="euk"] %.% sample_row(balanced_size),
    complete_dset_3mers_counts[class=="viruses"] %.% sample_row(balanced_size))



trainIndex <- createDataPartition(complete_dset_3mers_counts_bal$class, p = .6, list = FALSE, times = 1)

seqTrain <- complete_dset_3mers_counts_bal[ trainIndex[,1]][,c("class",count_cols_3mers),with=F]
seqTest  <- complete_dset_3mers_counts_bal[-trainIndex[,1],c("class",count_cols_3mers),with=F]

seqTrain_length=apply(seqTrain[,count_cols_3mers,with=F],1,sum)
seqTrain_length_normalized=seqTrain[,count_cols_3mers,with=F]/seqTrain_length
seqTrain_length_normalized$class=seqTrain$class

seqTest_length=apply(seqTest[,count_cols_3mers,with=F],1,sum)
seqTest_length_normalized=seqTest[,count_cols_3mers,with=F]/seqTest_length
seqTest_length_normalized$class=seqTest$class




svmTrainFit_3mers_balanced_complete <- train(class ~ ., data = seqTrain_length_normalized,
                 method = "svmRadial",
                 trControl = trainControl(method = "none"),
                 tuneGrid=data.frame(sigma=0.01,C=2),
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = T)

predictions=predict(svmTrainFit_3mers_balanced_complete,newdata=seqTest_length_normalized[,count_cols_3mers,with=F])
confusionMatrix(predictions,seqTest$class)
```
We get a slightly worse classifier: 

  Confusion Matrix and Statistics

            Reference
  Prediction archea bact  euk viruses
     archea    1073  135  143     155
     bact        69 1107   62     180
     euk        130   60 1003     149
     viruses    112   82  176     900

  Overall Statistics
                                            
                 Accuracy : 0.7375          
                   95% CI : (0.7257, 0.7491)
      No Information Rate : 0.25            
      P-Value [Acc > NIR] : < 2.2e-16       
                                            
                    Kappa : 0.65            
   Mcnemar's Test P-Value : 1.139e-12       

  Statistics by Class:

                       Class: archea Class: bact Class: euk Class: viruses
  Sensitivity                 0.7753      0.7999     0.7247         0.6503
  Specificity                 0.8957      0.9251     0.9184         0.9109
  Pos Pred Value              0.7125      0.7807     0.7474         0.7087
  Neg Pred Value              0.9228      0.9327     0.9092         0.8865
  Prevalence                  0.2500      0.2500     0.2500         0.2500
  Detection Rate              0.1938      0.2000     0.1812         0.1626
  Detection Prevalence        0.2720      0.2561     0.2424         0.2294
  Balanced Accuracy           0.8355      0.8625     0.8215         0.7806


# Classification using a very gapped spaced k-mer space 

```{r }

rm(list=ls())
gapped_4_spMers=fread("spaced_kmers/long_kmers_mVar_n30.fa_1011001.csv")

seq_attributes=limma::strsplit2(gapped_4_spMers$sequence_description,"_")
gapped_4_spMers$class=factor(seq_attributes[,2])
gapped_4_spMers$species=factor(seq_attributes[,1])
gapped_4_spMers$pattern=factor(gapped_4_spMers$pattern)
gapped_4_spMers$kmer=factor(gapped_4_spMers$kmer)
rm(seq_attributes)

gapped_4_spMers_counts=dcast.data.table(gapped_4_spMers,class+sequence_description~kmer,value.var='count',fill=0)
gapped_4_spMers_counts_cols=colnames(gapped_4_spMers_counts)[-(1:2)]
table(gapped_4_spMers_counts$class)

# We balance in terms of sequences 
sample_row = function(df,size){df[sample(1:nrow(df),size=size),]}
balanced_size=min(table(gapped_4_spMers_counts$class))

set.seed(3456)
gapped_4_spMers_counts_bal=rbind(
    gapped_4_spMers_counts[class=="archea"] %.% sample_row(balanced_size),
    gapped_4_spMers_counts[class=="bact"] %.% sample_row(balanced_size),
    gapped_4_spMers_counts[class=="euk"] %.% sample_row(balanced_size),
    gapped_4_spMers_counts[class=="viruses"] %.% sample_row(balanced_size))


trainIndex <- createDataPartition(gapped_4_spMers_counts_bal$class, p = .6, list = FALSE, times = 1)

seqTrain <- gapped_4_spMers_counts_bal[ trainIndex[,1]][,c("class",gapped_4_spMers_counts_cols),with=F]
seqTest  <- gapped_4_spMers_counts_bal[-trainIndex[,1],c("class",gapped_4_spMers_counts_cols),with=F]

seqTrain_length=apply(seqTrain[,gapped_4_spMers_counts_cols,with=F],1,sum)
seqTrain_length_normalized=seqTrain[,gapped_4_spMers_counts_cols,with=F]/seqTrain_length
seqTrain_length_normalized$class=seqTrain$class

seqTest_length=apply(seqTest[,gapped_4_spMers_counts_cols,with=F],1,sum)
seqTest_length_normalized=seqTest[,gapped_4_spMers_counts_cols,with=F]/seqTest_length
seqTest_length_normalized$class=seqTest$class


svm_gapped_4mers <- train(class ~ ., data = seqTrain_length_normalized,
                 method = "svmRadial",
                 trControl = trainControl(method = "none"),
                 tuneGrid=data.frame(sigma=0.01,C=2),
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = T)

predictions=predict(svm_gapped_4mers,newdata=seqTest_length_normalized[,gapped_4_spMers_counts_cols,with=F])
confusionMatrix(predictions,seqTest$class)
```

We get 

  Confusion Matrix and Statistics

            Reference
  Prediction archea bact  euk viruses
     archea    1050   93   96     123
     bact        81 1024   53     160
     euk        147  133 1068     194
     viruses    106  134  167     907

  Overall Statistics
                                           
                 Accuracy : 0.7314         
                   95% CI : (0.7195, 0.743)
      No Information Rate : 0.25           
      P-Value [Acc > NIR] : < 2.2e-16      
                                           
                    Kappa : 0.6419         
   Mcnemar's Test P-Value : 2.328e-09      

  Statistics by Class:

                       Class: archea Class: bact Class: euk Class: viruses
  Sensitivity                 0.7587      0.7399     0.7717         0.6553
  Specificity                 0.9249      0.9292     0.8858         0.9020
  Pos Pred Value              0.7709      0.7769     0.6926         0.6903
  Neg Pred Value              0.9200      0.9147     0.9209         0.8870
  Prevalence                  0.2500      0.2500     0.2500         0.2500
  Detection Rate              0.1897      0.1850     0.1929         0.1638
  Detection Prevalence        0.2460      0.2381     0.2785         0.2374
  Balanced Accuracy           0.8418      0.8345     0.8288         0.7787



# Gapped 5 mers 

```{r }
rm(list=ls())
gapped_5mers=fread("spaced_kmers/long_kmers_mVar_n30.fa_101001011.csv")

```


# Trying to sparsify count data 


# Very sparse count data 

```{r }
sparse_8mers=fread("spaced_kmers/long_kmers_mVar_n30.fa_1000010011001010101.csv")

seq_attributes=limma::strsplit2(sparse_8mers$sequence_description,"_")
sparse_8mers$class=factor(seq_attributes[,2])
sparse_8mers$species=factor(seq_attributes[,1])
sparse_8mers$pattern=factor(sparse_8mers$pattern)
sparse_8mers$kmer=factor(sparse_8mers$kmer)
rm(seq_attributes)


kmer_by_species=sparse_8mers[,sum(count),by=list(kmer,class,species)]
# We binarize 
kmer_by_species$V1=1

# This give the number of unique species in which we find a kmer 
kmer_by_unique_species=kmer_by_species[,sum(V1),by=list(kmer,class)][order(V1)]

kmer_by_unique_species_count=dcast.data.table(kmer_by_unique_species,kmer~class,fill=0)


n_classes=sparse_8mers[,.N,by=list(class,species)][,.N,by=class]
n_classes_l=as.list(n_classes$N)
names(n_classes_l)=as.character(n_classes$class)

```


We compute the chi2 

```{r }
kmer_viral_contingency=kmer_by_unique_species_count[,list(
  n_viral_no=n_classes_l$viruses - viruses, # N viral species without the k-mer 
  n_viral=viruses, # N viral species with the k-mer
  n_no_viral=archea+bact+euk, # N non-viral with the k-mer 
  n_no_viral_no=sum(unlist(n_classes_l))-n_classes_l$viruses - (archea+bact+euk) # N non-viral without the k-mer 
  ),by=kmer]

kmer_viral_chi2=kmer_viral_contingency[,list(viral_chi2=chisq.test(matrix(c(n_viral_no,n_viral,n_no_viral_no,n_no_viral),byrow=T,ncol=2))$statistic),by=kmer]


kmer_euk_contingency=kmer_by_unique_species_count[,list(
  n_euk_no=n_classes_l$euk - euk, # N euk species without the k-mer 
  n_euk=euk, # N euk species with the k-mer
  n_no_euk=archea+bact+viruses, # N non-euk with the k-mer 
  n_no_euk_no=sum(unlist(n_classes_l))-n_classes_l$euk - (archea+bact+viruses) # N non-euk without the k-mer 
  ),by=kmer]

kmer_euk_chi2=kmer_euk_contingency[,list(euk_chi2=chisq.test(matrix(c(n_euk_no,n_euk,n_no_euk_no,n_no_euk),byrow=T,ncol=2))$statistic),by=kmer]


kmer_bact_contingency=kmer_by_unique_species_count[,list(
  n_bact_no=n_classes_l$bact - bact, # N bact species without the k-mer 
  n_bact=bact, # N bact species with the k-mer
  n_no_bact=archea+bact+viruses, # N non-bact with the k-mer 
  n_no_bact_no=sum(unlist(n_classes_l))-n_classes_l$bact - (archea+euk+viruses) # N non-bact without the k-mer 
  ),by=kmer]

kmer_bact_chi2=kmer_bact_contingency[,list(bact_chi2=chisq.test(matrix(c(n_bact_no,n_bact,n_no_bact_no,n_no_bact),byrow=T,ncol=2))$statistic),by=kmer]


kmer_archea_contingency=kmer_by_unique_species_count[,list(
  n_archea_no=n_classes_l$archea - archea, # N archea species without the k-mer 
  n_archea=archea, # N archea species with the k-mer
  n_no_archea=archea+archea+viruses, # N non-archea with the k-mer 
  n_no_archea_no=sum(unlist(n_classes_l))-n_classes_l$archea - (archea+euk+viruses) # N non-archea without the k-mer 
  ),by=kmer]

kmer_archea_chi2=kmer_archea_contingency[,list(archea_chi2=chisq.test(matrix(c(n_archea_no,n_archea,n_no_archea_no,n_no_archea),byrow=T,ncol=2))$statistic),by=kmer]


kmer_all_contingency=kmer_by_unique_species_count[,list(
    n_archea_no=n_classes_l$archea - archea, # N archea species without the k-mer 
    n_bact_no=n_classes_l$bact - bact, # N archea species without the k-mer 
    n_euk_no=n_classes_l$euk - euk, # N archea species without the k-mer 
    n_viruses_no=n_classes_l$viruses - viruses, # N archea species without the k-mer 
    n_archea=archea, # N archea species without the k-mer 
    n_bact=bact, # N archea species without the k-mer 
    n_euk=euk, # N archea species without the k-mer 
    n_viruses=viruses # N archea species without the k-mer 
  ),by=kmer]

kmer_chi2=kmer_all_contingency[,list(global_chi2=chisq.test(matrix(c(n_archea_no, n_bact_no, n_euk_no, n_viruses_no, n_archea, n_bact, n_euk, n_viruses),byrow=T,ncol=4))$statistic),by=kmer]

kmer_by_unique_species_count=merge(merge(merge(merge(kmer_by_unique_species_count,kmer_viral_chi2,by="kmer"),kmer_euk_chi2,by="kmer"),kmer_bact_chi2,by='kmer'),kmer_archea_chi2,by="kmer")
kmer_by_unique_species_count=merge(kmer_by_unique_species_count,kmer_chi2,by='kmer')


#10.83 == signif at the 0.001 level a
g1=ggplot(kmer_by_unique_species_count,aes(x=archea+bact+euk,y=viruses,colour=interaction(global_chi2>=10.83,viral_chi2>=10.83)))+geom_point()+geom_abline(slope=1)
g2=ggplot(kmer_by_unique_species_count,aes(x=archea+bact+viruses,y=euk,colour=interaction(global_chi2>=10.83,euk_chi2>=10.83)))+geom_point()+geom_abline(slope=1)
g3=ggplot(kmer_by_unique_species_count,aes(x=archea+euk+viruses,y=bact,colour=interaction(global_chi2>=10.83,bact_chi2>=10.83)))+geom_point()+geom_abline(slope=1)
g4=ggplot(kmer_by_unique_species_count,aes(x=euk+bact+viruses,y=archea,colour=interaction(global_chi2>=10.83,archea_chi2>=10.83)))+geom_point()+geom_abline(slope=1)
grid.arrange(g1,g2,g3,g4)

```

If we combine the top-most viral specific k-mers, can we recover a significant fraction of viral species ?

```{r }
top_viral_kmers=tail(kmer_by_unique_species_count[order(viral_chi2),kmer],n=5)
sparse_8mers[kmer %in% top_viral_kmers][,length(unique(species)),by=list(class)]/n_classes$N

top_bact_kmers=tail(kmer_by_unique_species_count[order(bact_chi2),kmer],n=5)
sparse_8mers[kmer %in% top_bact_kmers][,length(unique(species)),by=list(class)]/n_classes$N

top_euk_kmers=tail(kmer_by_unique_species_count[order(euk_chi2),kmer],n=5)
sparse_8mers[kmer %in% top_euk_kmers][,length(unique(species)),by=list(class)]/n_classes$N

```