We test whether very long k-mers (e.g 10nt, starred or not) can be used as basis of separation between classes

```{r }
library(data.table)
library(caret)
library(reshape2)
library(party)
library(gridExtra)
library(ggplot2)
library(dplyr)
library(plyr)
library(corrplot)
setwd("/Users/hayssam/metaviro/experiments/140530_long_kmer_analysis")
```


```{r }
long_kmers=fread("long_kmers.fa.csv")
seq_attributes=limma::strsplit2(long_kmers$sequence_description,"_")
long_kmers$class=factor(seq_attributes[,2])
long_kmers$species=factor(seq_attributes[,1])
save(long_kmers,file="long_kmers.RData")
```

```{r}
load("long_kmers.RData")
```


Descriptive stats

```{r }
long_kmers[,.N,by=class]
```

K-mers by class ?

```{r }
long_kmers[,sum(count),by=list(kmer,class)][order(V1)] %.% tail(n=30)

```

How many ≠ species retrieved by a single k-mer ? 

```{r }
long_kmer_retrieval=long_kmers[,length(unique(species)),by=list(kmer,class)][order(V1)]
save(long_kmer_retrieval,file="long_kmer_retrieval.RData")

long_kmer_retrieval %.% tail(n=10)
```

Any viral specificity?

```{r}
long_kmer_retrieval_d = data.table(dcast(long_kmer_retrieval,kmer~class,value.var="V1",fill=0))

long_kmer_retrieval_d[order(viruses)]
long_kmer_retrieval_d[,purity:=viruses/(archea+bact+euk+viruses)]
long_kmer_retrieval_d[,euk_purity:=euk/(archea+bact+euk+viruses)]
long_kmer_retrieval_d[,bact_purity:=bact/(archea+bact+euk+viruses)]
long_kmer_retrieval_d[,arch_purity:=archea/(archea+bact+euk+viruses)]


long_kmer_retrieval_d[viruses>=10][order(purity)]

g1=ggplot(long_kmer_retrieval_d[viruses>=6],aes(x=viruses,y=purity,size=euk_purity,colour=arch_purity))+geom_point()+ylim(0,1)
g2=ggplot(long_kmer_retrieval_d[euk>=6],aes(x=euk,y=euk_purity,size=purity,colour=arch_purity))+geom_point()+ylim(0,1)
g3=ggplot(long_kmer_retrieval_d[bact>=6],aes(x=bact,y=bact_purity,size=purity,colour=arch_purity))+geom_point()+ylim(0,1)
g4=ggplot(long_kmer_retrieval_d[archea>=6],aes(x=archea,y=arch_purity,size=purity,colour=euk_purity))+geom_point()+ylim(0,1)
grid.arrange(g1,g2,g3,g4)
ggsave("comparative_purities.pdf")

long_kmer_retrieval_d[,virus_recall:=viruses/30]
long_kmer_retrieval_d[,virus_precision:=purity]
long_kmer_retrieval_d[,virus_f1_score:=virus_recall*virus_precision/(virus_precision+virus_recall)]

long_kmer_retrieval_d[,euk_recall:=euk/30]
long_kmer_retrieval_d[,euk_precision:=euk_purity]
long_kmer_retrieval_d[,euk_f1_score:=euk_recall*euk_precision/(euk_precision+euk_recall)]

long_kmer_retrieval_d[,bact_recall:=bact/30]
long_kmer_retrieval_d[,bact_precision:=bact_purity]
long_kmer_retrieval_d[,bact_f1_score:=bact_recall*bact_precision/(bact_precision+bact_recall)]

long_kmer_retrieval_d[,arch_recall:=archea/30]
long_kmer_retrieval_d[,arch_precision:=arch_purity]
long_kmer_retrieval_d[,arch_f1_score:=arch_recall*arch_precision/(arch_precision+arch_recall)]

ggplot(long_kmer_retrieval_d[viruses>=5],aes(x=virus_precision,y=virus_recall,colour=virus_f1_score))+geom_point()+xlim(0,1)+ylim(0,1)
ggsave("virus_prec_recall.pdf")

ggplot(long_kmer_retrieval_d[euk>=5],aes(x=euk_precision,y=euk_recall,colour=euk_f1_score))+geom_point()+xlim(0,1)+ylim(0,1)

long_kmer_retrieval_d[!is.nan(euk_f1_score)][order(euk_f1_score)] %.% tail(n=3)
```

We select 3 champions per class

```{r }
champions=unique(c(
	(long_kmer_retrieval_d[!is.nan(virus_f1_score)][order(virus_f1_score)] %.% tail(n=30))$kmer,
	(long_kmer_retrieval_d[!is.nan(bact_f1_score)][order(bact_f1_score)] %.% tail(n=30))$kmer,
	(long_kmer_retrieval_d[!is.nan(arch_f1_score)][order(arch_f1_score)] %.% tail(n=30))$kmer,
	(long_kmer_retrieval_d[!is.nan(euk_f1_score)][order(euk_f1_score)] %.% tail(n=30))$kmer
	))
long_kmers[kmer %in% champions,.N,by=list(class,species)][,length(unique(species)),by=class]

table(long_kmers[kmer %in% champions,class])

champions_counts=dcast(long_kmers[kmer %in% champions],class+species~kmer,fun.aggregate=sum,value.var="count")
champions_counts$species<-NULL

tt=ctree(data=champions_counts,class~.)
plot(tt)

confusionMatrix(predict(tt,champions_counts),champions_counts$class)
```

We should seek a combination of k-mer maximizing the overall F1-score of the viral classes 

We check whether this holds on a indepdendant subset 

```{r }
long_kmers_b2=fread("long_kmers_b2.fa.csv")
seq_attributes=limma::strsplit2(long_kmers_b2$sequence_description,"_")
long_kmers_b2$class=factor(seq_attributes[,2])
long_kmers_b2$species=factor(seq_attributes[,1])
save(long_kmers_b2,file="long_kmers_b2.RData")
long_kmers_b2[,.N,by=list(class,species)][,length(unique(species)),by=class]
```

We make the count matrix with the previous champions

```{r }
champions_counts_b2=dcast(long_kmers_b2[kmer %in% champions],class+species~kmer,fun.aggregate=sum,value.var="count")
champions_counts_b2$species<-NULL
tt2=ctree(data=champions_counts_b2,class~.)
plot(tt2)


confusionMatrix(predict(tt,champions_counts_b2),champions_counts_b2$class)
confusionMatrix(predict(tt2,champions_counts),champions_counts$class)
```