We want to determine whether viruses are more spread than the other classes

```{r libraries}
library(data.table)
library(cluster)
library(ggplot2)
library(corrplot)
library(plyr)
library(dplyr)
library(reshape2)
library(fpc)
library(foreach)
library(doParallel)
registerDoParallel(cores=2)


setwd("~/Documents/metaviro/experiments/141003_GenBank_resampling/")

apply(expand.grid("500nt_Resample_",1:50,"/",c("archaea","bact"),"_k3_mers.tsv"),FUN=function(x) paste(x,collapse="",sep=""),MARGIN=1)
pb <- txtProgressBar(min = 0, max = 50, style = 3)

all_instances_kmers= foreach(res_id=1:50,.combine=rbind) %do% {
	setTxtProgressBar(pb, res_id)
	rbind(
	fread(paste0("500nt_Resample_",res_id,"/archaea_k3_mers.tsv")),
	fread(paste0("500nt_Resample_",res_id,"/bact_k3_mers.tsv")),
	fread(paste0("500nt_Resample_",res_id,"/euk_k3_mers.tsv")),
	fread(paste0("500nt_Resample_",res_id,"/viruses_k3_mers.tsv")))
}
close(pb)

# let's annotate the viruses 


kmers_columns=colnames(all_instances_kmers)[5:ncol(all_instances_kmers)]

sequence_attributes=data.table(cbind(
	limma::strsplit2(all_instances_kmers$path,"_|/",fixed=F),
	limma::strsplit2(all_instances_kmers$sequence_description,"|",fixed=T)))

sequence_attributes[,contig_index:=1:.N,by=V4]


all_instances_kmers$gi=as.integer(as.character(sequence_attributes[,V7]))
all_instances_kmers$domain=factor(sequence_attributes[,V4])
all_instances_kmers$batch=factor(sequence_attributes[,V2])
all_instances_kmers$mean_length=factor(sequence_attributes[,V3])
all_instances_kmers$contig_index=sequence_attributes$contig_index

load("gi_mapping/viruses_annotations_gi.RData")
GenBank_annotations[,.N,by=Host][order(N)]
annotated_kmers=merge(all_instances_kmers,GenBank_taxID2gi,by="gi",all.x=T)
annotated_kmers=merge(annotated_kmers,GenBank_annotations,by="TaxID",all.x=T)

kmer_tot=rowSums(annotated_kmers[,kmers_columns,with=F])
annotated_kmers$kmer_tot=kmer_tot
annotated_kmers[,ratio:=kmer_tot/sequence_length,by=.I]

annotated_kmers[order(sequence_length)]
annotated_kmers=annotated_kmers[sequence_length<=2000]
annotated_kmers=annotated_kmers[sequence_length>=150]
annotated_kmers=annotated_kmers[ratio>=0.6]


summary(annotated_kmers$sequence_length)
summary(annotated_kmers$kmer_tot)
summary(annotated_kmers$ratio)

ggplot(annotated_kmers[sample(nrow(annotated_kmers),1e5)],aes(x=sequence_length,y=ratio))+geom_point()


# normalize and type 
normalized_kmers_counts=annotated_kmers[,kmers_columns,with=F]/annotated_kmers$sequence_length

normalized_kmers=cbind(normalized_kmers_counts, annotated_kmers[,list(domain,sequence_length,gi,BioProject.ID,Group,SubGroup,"tgt_host"=Host,"org"=Organism.Name,"org.GC"=GC.,"size"=Size,"seq.GC"=GC)] )

normalized_kmers$org.GC=as.numeric(normalized_kmers$org.GC)
normalized_kmers$size=as.numeric(normalized_kmers$size)

# normalized_kmers$tgt_host=as.factor(normalized_kmers$tgt_host)
normalized_kmers$Group=as.factor(normalized_kmers$Group)
normalized_kmers$SubGroup=as.factor(normalized_kmers$SubGroup)
normalized_kmers$gi=as.factor(normalized_kmers$gi)
normalized_kmers$org=as.factor(normalized_kmers$org)
normalized_kmers$idx=1:nrow(normalized_kmers)
normalized_kmers$classLabels<-NULL
normalized_kmers[,classLabels:=as.character(ifelse(is.na(tgt_host),ifelse(domain=="viruses","viruses_unassigned",as.character(domain)),paste("viruses",as.character(tgt_host),sep="_")))]
# normalized_kmers[grep("plants",classLabels),classLabels]
sort(table(normalized_kmers$classLabels))
# save(normalized_kmers,file="combined_resample_1_50_normalized_kmers.RData")


```

* We remove forbidden GIs 

```{r }
bad_gis=fread("blast_dbs/bad_gis.txt")$V2
normalized_kmers=normalized_kmers[!(gi %in% bad_gis)]

```


* We balance downsample 

```{r }
# most_frequent_classes=tail(normalized_kmers[grep("unclassified|unassigned|unknown",classLabels,invert=T)][,.N,by=classLabels][order(N)],n=8)[,classLabels]
most_frequent_classes=c("viruses_archaea","viruses_bacteria","viruses_plants","bact","archaea","euk")

# viruses_and_hosts=normalized_kmers[as.character(classLabels) %in% most_frequent_classes]
# save(viruses_and_hosts,file="combined_resample_1_50_viruses_and_hosts.RData")

min_size=min(7000,normalized_kmers[as.character(classLabels) %in% most_frequent_classes,.N,by=classLabels][,min(N)])

# normalized_kmers[,.N,by=classLabels][order(N)]
downsample_idx=unlist(llply(most_frequent_classes,function(g){sample(normalized_kmers[as.character(classLabels) == g,idx],min_size)}))

normalized_kmers_ds_SG=normalized_kmers[idx %in% downsample_idx]

normalized_kmers_ds_SG[,.N,by=classLabels]
dim(normalized_kmers_ds_SG)
```

* MCA analysis 

```{r }
library(FactoMineR)
library(broom)

res.mca=PCA(normalized_kmers_ds_SG[,c("classLabels","seq.GC",kmers_columns),with=F],quali.sup=c(1),quanti.sup=c(2),graph=F,scale.unit=T)
foo=dimdesc(res.mca)



# plot.PCA(res.mca,invisible=c("quali"),habillage=c(1),label="none")
plotellipses(res.mca,invisible=c("quali"),habillage=c(1),label="none")

plot.PCA(res.mca,invisible=c("quali"),habillage=c(1),axes=c(1,3),label="none")

plot.PCA(res.mca,invisible=c("quali"),habillage=c(1),axes=c(2,3),label="none")
plotellipses(res.mca,invisible=c("quali"),habillage=c(1),axis=c(1,3),label="none")

plotellipses(res.mca,invisible=c("quali"),habillage=c(1),axis=c(2,3),label="none")

plotellipses(res.mca,invisible=c("quali"),habillage=c(1),axis=c(3,5),label="none")

plot.PCA(res.mca,invisible=c("quali"),habillage=c(1),axes=c(2,3),label="none")
dimdesc(res.mca)

```
* we make nicer plots with ggplots 

```{r }
library(ggsubplot)

res.mca=PCA(normalized_kmers_ds_SG[,c("classLabels","seq.GC",kmers_columns),with=F],quali.sup=c(1),quanti.sup=c(2),graph=F,scale.unit=T)

ind_coord=data.table(res.mca$ind$coord)
ind_coord$classLabels=normalized_kmers_ds_SG$classLabels
ind_coord$domain=normalized_kmers_ds_SG$domain
ind_coord$seq.GC=normalized_kmers_ds_SG$seq.GC
ind_coord$top_level_group=normalized_kmers_ds_SG$Group
ind_coord$sub_group=normalized_kmers_ds_SG$SubGroup

sub_group_to_keep=names(tail(sort(table(ind_coord$sub_group)),n=10))
ind_coord[domain=="viruses" & !(sub_group %in% sub_group_to_keep),sub_group:="misc."]
ind_coord[domain=="viruses" & sub_group=="unclassified",sub_group:="misc."]
sort(table(ind_coord$sub_group))

ind_coord[,top_level_domain:="Cellular"]
ind_coord[domain=="viruses",top_level_domain:="Viral"]

correct_names=list("viruses_archaea")



# ggplot(ind_coord,aes(x=Dim.1,y=Dim.3,colour=classLabels,fill=classLabels))+geom_point()+stat_binhex(alpha=0.1,bins=10)
# ggplot(ind_coord,aes(x=Dim.1,y=Dim.3,colour=classLabels,fill=classLabels))+geom_point()+stat_bin2d(alpha=0.1)

# ggplot(ind_coord,aes(x=Dim.1,y=Dim.2,colour=classLabels,fill=classLabels))+geom_point()+geom_density2d()+facet_wrap(~domain)

ggplot(ind_coord,aes(x=Dim.1,y=Dim.2,colour=classLabels,fill=classLabels))+geom_point()+geom_density2d()+facet_wrap(~classLabels)

ggplot(ind_coord,aes(x=Dim.1,y=Dim.2,colour=seq.GC))+geom_point()+geom_density2d()+facet_wrap(~classLabels)

ggplot(ind_coord,aes(x=Dim.1,y=Dim.2,colour=sub_group,group=classLabels))+geom_point(alpha=0.5)+geom_density2d(colour="firebrick2",linemitre=1)+facet_wrap(~classLabels)

ggplot(ind_coord,aes(x=Dim.1,y=Dim.2,colour=sub_group,group=classLabels))+geom_point(alpha=0.5)+geom_density2d(colour="firebrick2",linemitre=1)+facet_wrap(~classLabels)+xlim(c(-12,12))+ylim(c(-5.5,5.5))
ggplot(ind_coord,aes(x=Dim.1,y=Dim.3,colour=sub_group,group=classLabels))+geom_point(alpha=0.5)+geom_density2d(colour="firebrick2",linemitre=1)+facet_wrap(~classLabels)+xlim(c(-12,12))+ylim(c(-5.5,5.5))
ggplot(ind_coord,aes(x=Dim.2,y=Dim.3,colour=sub_group,group=classLabels))+geom_point(alpha=0.5)+geom_density2d(colour="firebrick2",linemitre=1)+facet_wrap(~classLabels)+xlim(c(-5.5,5.5))+ylim(c(-5.5,5.5))
# ggplot(ind_coord,aes(x=Dim.1,y=Dim.2,group=classLabels))+geom_point(alpha=0.5)+stat_density2d(aes(fill = ..level..),geom="polygon")+facet_wrap(~classLabels)+xlim(c(-12,12))+ylim(c(-5.5,5.5))

ggplot(ind_coord,aes(x=Dim.1,y=Dim.3,colour=sub_group,group=classLabels))+geom_point(alpha=0.5)+geom_density2d(colour="firebrick2")+facet_wrap(~classLabels)+xlim(c(-12,12))+ylim(c(-5.5,5.5))
ggplot(ind_coord,aes(x=Dim.2,y=Dim.3,colour=sub_group,group=classLabels))+geom_point(alpha=0.5)+geom_density2d(colour="firebrick2")+facet_wrap(~classLabels)+xlim(c(-10,10))+ylim(c(-10,10))

ggplot(ind_coord,aes(x=Dim.1,y=Dim.3,colour=classLabels))+geom_point()+geom_density2d()+facet_wrap(~classLabels)

ggplot(ind_coord,aes(x=Dim.1,y=Dim.3,colour=classLabels,fill=classLabels))+geom_point()+geom_density2d()+facet_wrap(~classLabels)
ggplot(ind_coord,aes(x=Dim.2,y=Dim.3,colour=classLabels,fill=classLabels))+geom_point()+geom_density2d()+facet_wrap(~classLabels)

# ggplot(ind_coord,aes(x=Dim.1,y=Dim.3,colour=classLabels,fill=classLabels))+geom_point()+geom_density2d()+facet_wrap(~domain)

```

* We determine the highest contours 

```{r }
ind_kde_arch=kde2d(ind_coord[classLabels=="archaea",Dim.1],ind_coord[classLabels=="archaea",Dim.2])

ind_kde_arch=kde2d(ind_coord[classLabels=="viruses_archaea",Dim.1],ind_coord[classLabels=="viruses_archaea",Dim.2],n=30)

coords_max=which(ind_kde_arch$z>=quantile(ind_kde_arch$z,0.90),arr.ind=T)

medoids_kde=pam(coords_max,k=3)
coords_max_med=medoids_kde$medoids

ind_kde_arch$x[coords_max_med[,1]]
ind_kde_arch$y[coords_max_med[,2]]


plot(ind_coord[classLabels=="viruses_archaea",Dim.1],ind_coord[classLabels=="viruses_archaea",Dim.2],xlim=c(-10,10),ylim=c(-6,6))
points(ind_kde_arch$x[coords_max_med[,1]],ind_kde_arch$y[coords_max_med[,2]],col="red",cex=4,pch='O')
points(ind_kde_arch$x[coords_max[,1]],ind_kde_arch$y[coords_max[,2]],col="red",cex=4)

```

* Using EMD 
Works perfectly well to find the local maxima in a 2d kernel 

```{r }
library(EMD)

tgt_class_label="viruses_archaea"
local_maximas=data.table(ldply(unique(ind_coord$classLabels),function(cl){
	ind_kde_local=kde2d(ind_coord[classLabels==cl,Dim.1],ind_coord[classLabels==cl,Dim.2],n=30)

	ind_kde_local$z[ind_kde_local$z<=quantile(ind_kde_local$z,0.92)]<-0
	ex=extrema2dC(ind_kde_local$z)
	max_ind=ind_kde_local$z
	max_ind[ex$maxindex]=1
	coords_max_med=which(max_ind==1,arr.ind=T)


	# plot(ind_coord[classLabels==tgt_class_label,Dim.1],ind_coord[classLabels==tgt_class_label,Dim.2],xlim=c(-10,10),ylim=c(-6,6))
	# points(ind_kde_local$x[coords_max_med[,1]],ind_kde_local$y[coords_max_med[,2]],col="red",cex=4,pch='O')
	data.frame(Dim.1=ind_kde_local$x[coords_max_med[,1]],Dim.2=ind_kde_local$y[coords_max_med[,2]],classLabels=cl,sub_group=NA,size=ind_kde_local$z[coords_max_med])
}))



local_maximas[,shape:=ifelse(grepl("virus",classLabels),21,22)]

save(local_maximas,file="local_maximas.RData")

```

* We visualize dist between local maximas

```{r }
library(corrplot)
local_maximas$idx=1:nrow(local_maximas)
g2=ggplot(local_maximas,aes(x=Dim.1,y=Dim.2,fill=classLabels,label=classLabels,shape=factor(shape)))+geom_point(size=12)+geom_text(hjust=0.5,vjust=2.5)+scale_shape_manual(values=c(21,22))+geom_text(aes(label=idx))
g2
local_maximas_mat=as.matrix(local_maximas[,list(Dim.1,Dim.2)])
rownames(local_maximas_mat)=make.names(local_maximas$classLabels,unique=T)

local_maximas_d=dist(local_maximas_mat)
corrplot(as.matrix(local_maximas_d),is.corr=F,order="hclust")

```
* We add kmeans 

```{r }
# pam(ind_coord[,list(Dim.1,Dim.2)],k=3)




clust_per_class=list("archaea"=3,"bact"=3,"euk"=1,"viruses_archaea"=3,"viruses_bacteria"=3,'viruses_plants'=1)


inds_kmeans=data.table(ldply(unique(ind_coord$classLabels),function(g){
	kmeans_clusts=kmeans(ind_coord[classLabels==g,list(Dim.1,Dim.2)],centers=clust_per_class[[g]]);
	data.table(kmeans_clusts$centers,classLabels=g,sub_group=NA,size=kmeans_clusts$size)
}))

inds_kmeans[,shape:=ifelse(grepl("virus",classLabels),21,22)]


### UGLY UGLY UGLY, make a simple hash table !
class_mapping=list("viruses_archaea"="archaea","viruses_bacteria"="bact","viruses_plants"="euk")
mapped_maximas=local_maximas[grep("viruses",classLabels),list(classLabels,Dim.1,Dim.2,sub_group,size,shape=21)]
mapped_maximas[,classLabels:=limma::strsplit2(classLabels,"_")[,2],by=.I]
mapped_maximas[classLabels=="plants",classLabels:="euk"]
mapped_maximas[classLabels=="bacteria",classLabels:="bact"]

inds_maximas=rbind(local_maximas,mapped_maximas)



# Clean up for panel names 


panel_names <- list(
	"archaea"="Archaea",
	"bact"="Bacteria",
	"euk"="Plants & Fungi",
	"viruses_archaea"="Archaeal viruses",
	"viruses_bacteria"="Bacterial viruses",
	"viruses_plants"="Plants viruses")

# names(panel_names)
ind_coord$classLabels=factor(ind_coord$classLabels,levels=names(panel_names),labels=as.character(panel_names))
inds_maximas$classLabels=factor(inds_maximas$classLabels,levels=names(panel_names),labels=as.character(panel_names))

# inds_pams=ldply(unique(ind_coord$classLabels),function(g)data.table(pam(ind_coord[classLabels==g,list(Dim.1,Dim.2)],k=clust_per_class[g])$medoids,classLabels=g,sub_group=NA,shapes=g))

# ggplot(ind_coord,aes(x=Dim.1,y=Dim.2,colour=classLabels))+geom_point(alpha=0.4)+guides(fill=FALSE,shape=FALSE) +xlim(c(-12,12))+ylim(c(-5.5,5.5))+ theme(legend.position="top") 


# ggplot(ind_coord,aes(x=Dim.1,y=Dim.2,colour=sub_group,group=classLabels))+geom_point(alpha=0.5)+geom_density2d(colour="firebrick2",linemitre=1)+facet_wrap(~classLabels)+geom_point(data=inds_maximas,shape=22,aes(size=size),fill="white")+ guides(fill=FALSE,colour=FALSE,shape=F) +xlim(c(-12,12))+ylim(c(-5.5,5.5))+scale_size_continuous(range=c(4,8))

g=ggplot(ind_coord,aes(x=Dim.1,y=Dim.2,colour=sub_group,group=classLabels))+geom_point(alpha=0.5)+geom_density2d(colour="firebrick2",linemitre=1)+facet_wrap(~classLabels)+xlim(c(-12,12))+ylim(c(-5.5,5.5))+scale_size_continuous(range=c(4,8))

g=g+geom_point(data=inds_maximas,aes(size=size,shape=factor(shape)),fill="white")+scale_shape_manual(values=c(21,22),name="Local density maxima:",labels=c("Host","Virus"))+guides(fill=FALSE,size=FALSE)+ theme_grey(base_size=22)+theme(legend.position="top",axis.title.x = element_blank(),axis.title.y= element_blank())+scale_colour_discrete(name="Viral sub-groups")
g
ggsave("PCA_plots_of_viral_and_hosts_v2.pdf",w=14,h=8)


# g+geom_point(data=inds_kmeans[grep("virus",classLabels,invert=F)],shape=22,aes(size=size),fill="white")
# We report the viruses contigs centers on the host space 

# g+geom_point(data=mapped_kmeans,shape=22,aes(size=size),fill="white")


# ggplot(ind_coord,aes(x=Dim.1,y=Dim.2,colour=sub_group,group=classLabels))+geom_point(alpha=0.5)+geom_density2d(colour="firebrick2",linemitre=1)+facet_wrap(~classLabels)+geom_point(data=inds_pams,aes(shape=shapes),size=5,fill="black",colour="black")+ guides(fill=FALSE,colour=FALSE,shape=FALSE)+xlim(c(-12,12))+ylim(c(-5.5,5.5))


```

we combine 

```{r }
library(gridExtra)
grid.arrange(g,g2)
g2

```

Other versions without mapped maximas but with labels 


```{r }

local_maximas_renamed=local_maximas
local_maximas_renamed$classLabels=factor(local_maximas_renamed$classLabels,levels=names(panel_names),labels=as.character(panel_names))

local_maximas_renamed=local_maximas_renamed[idx!=9]

local_maximas_renamed[order(classLabels),idx:=1:nrow(local_maximas_renamed)]

g=ggplot(ind_coord,aes(x=Dim.1,y=Dim.2,colour=sub_group,group=classLabels))+geom_point(alpha=0.5)+geom_density2d(colour="firebrick2",linemitre=1)+facet_wrap(~classLabels)+xlim(c(-12,12))+ylim(c(-5.5,5.5))+scale_size_continuous(range=c(4,8))
g=g+geom_point(data=local_maximas_renamed,aes(size=size,shape=factor(shape)),fill="white")+scale_shape_manual(values=c(21,22),name="Local density maxima:",labels=c("Host","Virus"))+guides(fill=FALSE,size=FALSE)+ theme_grey(base_size=22)+theme(legend.position="top",axis.title.x = element_blank(),axis.title.y= element_blank())+scale_colour_discrete(name="Viral sub-groups")
g=g+geom_text(data=local_maximas_renamed,aes(label=idx),colour="black")+scale_size_continuous(range=c(8,12))
g


# make g2 for local maximas 

# Adjust labels 

local_maximas_renamed$label.x=local_maximas_renamed$Dim.1
local_maximas_renamed$label.y=local_maximas_renamed$Dim.2
local_maximas_renamed[idx==12,label.y:=label.y+0.35]
local_maximas_renamed[idx==7,label.y:=label.y+0.35]
local_maximas_renamed[idx==8,label.y:=label.y+0.35]

g2=ggplot(local_maximas_renamed,aes(x=Dim.1,y=Dim.2,fill=classLabels,shape=factor(shape)))+geom_point(size=12)+geom_text(aes(label=classLabels,x=label.x,y=label.y),hjust=0.5,vjust=2.5)+scale_shape_manual(values=c(21,22))+geom_text(aes(label=idx))+guides(fill=FALSE,size=FALSE,shape=FALSE)+theme(axis.title.x = element_blank(),axis.title.y= element_blank())
g2


grid.arrange(g,g2)
g2



```

# Using and MDS 
Too costly !

```{r }

min_size=min(1500,normalized_kmers[as.character(classLabels) %in% most_frequent_classes,.N,by=classLabels][,min(N)])

normalized_kmers[,.N,by=classLabels][order(N)]
downsample_idx=unlist(llply(most_frequent_classes,function(g){sample(normalized_kmers[as.character(classLabels) == g,idx],min_size)}))


normalized_kmers_ds_SG_1500=normalized_kmers[downsample_idx]
normalized_kmers_ds_SG_1500[,.N,by=classLabels]
dim(normalized_kmers_ds_SG_1500)


d <- dist(normalized_kmers_ds_SG_1500[,kmers_columns,with=F])
fit <- cmdscale(d,eig=TRUE, k=2) # k is the number of dim
str(fit)

mds_coords=data.table(fit$points)

mds_coords$classLabels=normalized_kmers_ds_SG_1500$classLabels
mds_coords$domain=normalized_kmers_ds_SG_1500$domain
mds_coords$seq.GC=normalized_kmers_ds_SG_1500$seq.GC
mds_coords$top_level_group=normalized_kmers_ds_SG_1500$Group
mds_coords$sub_group=normalized_kmers_ds_SG_1500$SubGroup
mds_coords$gi=normalized_kmers_ds_SG_1500$gi
mds_coords$idx=normalized_kmers_ds_SG_1500$idx
save(mds_coords,file="1500_per_goupds_mds_coords.RData")

# test pam again 

# pamk(mds_coords[classLabels=="bact",list(V1,V2)])

# mds_pams=pam(mds_coords[classLabels=="bact",list(V1,V2)],k=2)
# mds_pams$medoids

# mds_centers=ldply(unique(mds_coords$classLabels),function(g)data.table(kmeans(mds_coords[classLabels==g,list(V1,V2)],centers=2)$centers,classLabels=g,sub_group=NA,shapes=g))
clust_per_class=list("archaea"=2,"bact"=2,"euk"=2,"viruses_archaea"=2,"viruses_bacteria"=2,'viruses_plants'=1)

mds_pams=ldply(unique(mds_coords$classLabels),function(g)data.table(pam(mds_coords[classLabels==g,list(V1,V2)],k=clust_per_class[g])$medoids,classLabels=g,sub_group=NA,shapes=g))

# We manually add some "attractors"
mds_centers=data.table(
	V1=rep(c(0.09,-0.037,0.075),each=6),
	V2=rep(c(0.005,0.01,0.005),each=6),
	shapes=rep(c("A","E","B"),each=6),
	classLabels=rep(c("archaea","euk","bact","viruses_bacteria","viruses_plants","viruses_archaea"),3),
	sub_group=NA
	)


# mds_coords$classLabels=factor(mds_coords$classLabels,levels=c("archaea","viruses_archaea","bact","viruses_bacteria","euk","viruses_plants"))
# mds_pams$classLabels=factor(mds_pams$classLabels,levels=c("archaea","viruses_archaea","bact","viruses_bacteria","euk","viruses_plants"))

vline_matching=mds_pams
# We dupl to make lines between viruses 

ggplot(mds_coords,aes(x=V1,y=V2,colour=sub_group,group=classLabels))+geom_point(alpha=0.5)+geom_density2d(colour="firebrick2",linemitre=1)+facet_wrap(~classLabels)+geom_point(data=mds_centers,aes(shape=shapes),size=5,fill="black",colour="black")+ guides(fill=FALSE,colour=FALSE)

ggplot(mds_coords,aes(x=V1,y=V2,colour=sub_group,group=classLabels))+geom_point(alpha=0.5)+geom_density2d(colour="firebrick2",linemitre=1)+facet_wrap(~classLabels)+geom_point(data=mds_pams,aes(shape=shapes),size=5,fill="black",colour="black")+ guides(fill=FALSE,colour=FALSE,shape=FALSE)+xlim(c(-0.14,0.14))+ylim(c(-0.06,0.06))

```


