We start with a fixed number of randomly sampled contigs for each input sequences.


```{r loading}
setwd("~/Documents/metaviro/experiments/140529_contig_variability")
library(data.table)
library(ggplot2)
library(reshape2)
kmers=fread("full_ncbi_contigs_per_sequences_3mer.csv")
seq_attributes=limma::strsplit2(kmers$sequence_description,"_")
kmers$class=seq_attributes[,2]
kmers$species=seq_attributes[,1]
kmer_columns=colnames(kmers)[5:(ncol(kmers)-2)]
```

How many ≠ species per class? 

```{r }
kmers[,.N,by=list(species,class)][,.N,by=class]

```


We plot the distribution for a given species 

```{r }

a_bact_species=kmers[species==kmers[class=="bact"][1,species]]
rowSums(a_bact_species[,kmer_columns,with=F])
kmers_melted=melt(a_bact_species[,c("sequence_description","sequence_length", kmer_columns),with=F],id.vars=c("sequence_description","sequence_length"))
ggplot(kmers_melted,aes(x=variable,y=value,color=sequence_description,group=sequence_description))+geom_point()+geom_line()
```

We determine the correlation between 10 bact and 10 arch species 

```{r }
selected_species=c(kmers[class=="virus"][1:5,species],kmers[class=="arch"][1:5,species])
sub_sample_kmers=kmers[species %in% selected_species]
kmers_melted=melt(sub_sample_kmers[,c("sequence_description","class","species","sequence_length",kmer_columns),with=F],id.vars=c("sequence_description","class","species","sequence_length"))
# We normalize 
kmers_melted[,value:=value/sequence_length]
ggplot(kmers_melted,aes(x=variable,y=value,color=class,group=sequence_description))+geom_point()+geom_line()+facet_wrap(~species,ncol=1,scale="free_y")
```

We do a permutation of the species 

```{r }
kmers_melted_permuted=kmers_melted
permutation=sample(nrow(kmers_melted_permuted))
kmers_melted_permuted$value = kmers_melted[permutation,value]
ggplot(kmers_melted_permuted,aes(x=variable,y=value,color=class,group=sequence_description))+geom_point()+geom_line()+facet_wrap(~species,ncol=1,scale="free_y")


```

We compare for a given species

```{r }
library(gridExtra)
g1=ggplot(kmers_melted[species=="162329288"],aes(x=variable,y=value,color=class,group=sequence_description))+geom_point()+facet_wrap(~species,ncol=1,scale="free_y")+ylim(0,0.1)+geom_line()
g2=ggplot(kmers_melted_permuted[species=="162329288"],aes(x=variable,y=value,color=class,group=sequence_description))+geom_point()+facet_wrap(~species,ncol=1,scale="free_y")+ylim(0,0.1)+geom_line()
grid.arrange(g1,g2)

```

We plot the correlation 

```{r }
library(corrplot)
sub_sample_kmers_df=data.frame(sub_sample_kmers[,kmer_columns,with=F],row.names=sub_sample_kmers$sequence_description)[1:40,]
corrplot(cor(t(sub_sample_kmers_df)),order="hclust")


```

We plot the average kmer freq for each sequence (thus across 10 contigs)

```{r }
ggplot(kmers_melted,aes(x=variable,y=value,color=class))+geom_boxplot()

```

For all samples

```{r }
selected_species=c(kmers[class=="virus",species],kmers[class=="arch",species],kmers[class=="bact",species],kmers[class=="euk",species])
sub_sample_kmers=kmers[species %in% selected_species]
kmers_melted=melt(sub_sample_kmers[,c("sequence_description","class","species","sequence_length",kmer_columns),with=F],id.vars=c("sequence_description","class","species","sequence_length"))
kmers_melted$value <- kmers_melted$value / kmers_melted$sequence_length

kmers_melted$variable_f <- as.numeric(factor(kmers_melted$variable))
# ggplot(kmers_melted[variable_f<=12],aes(x=factor(variable),y=value,fill=class))+geom_boxplot()+ylim(0,50)
# ggplot(kmers_melted[12<variable_f & variable_f<=24],aes(x=factor(variable),y=value,fill=class))+geom_boxplot()+ylim(0,50)
# ggplot(kmers_melted[24<variable_f & variable_f<=36],aes(x=factor(variable),y=value,fill=class))+geom_boxplot()+ylim(0,0.1)
# ggplot(kmers_melted[36<variable_f & variable_f<=48],aes(x=factor(variable),y=value,fill=class))+geom_boxplot()+ylim(0,0.1)
ggplot(kmers_melted[48<variable_f & variable_f<=64],aes(x=factor(variable),y=value,fill=class))+geom_boxplot()+ylim(0,0.1)


```

Correlation between k-mers by classes 

```{r }
sub_sample_kmers_df=data.frame(kmers[,kmer_columns,with=F],row.names=kmers$sequence_description)
kmer_cor=cor(sub_sample_kmers_df)

kmer_order = corrplot(kmer_cor,order="hclust")
kmer_order=colnames(kmer_order)

```

We do it class by class

```{r }
sub_sample_kmers_df_bact=data.frame(kmers[class=="bact"][,kmer_columns,with=F],row.names=kmers[class=="bact"]$sequence_description)
sub_sample_kmers_df_virus=data.frame(kmers[class=="virus"][,kmer_columns,with=F],row.names=kmers[class=="virus"]$sequence_description)
sub_sample_kmers_df_euk=data.frame(kmers[class=="euk"][,kmer_columns,with=F],row.names=kmers[class=="euk"]$sequence_description)
sub_sample_kmers_df_arch=data.frame(kmers[class=="arch"][,kmer_columns,with=F],row.names=kmers[class=="arch"]$sequence_description)

kmer_cor_bact=cor(sub_sample_kmers_df_bact)[kmer_order,kmer_order]
kmer_cor_virus=cor(sub_sample_kmers_df_virus)[kmer_order,kmer_order]
kmer_cor_euk=cor(sub_sample_kmers_df_euk)[kmer_order,kmer_order]
kmer_cor_arch=cor(sub_sample_kmers_df_arch)[kmer_order,kmer_order]

par(mfrow=c(2,2))
corrplot(kmer_cor_bact,main="bact")
corrplot(kmer_cor_virus,main="virus")
corrplot(kmer_cor_euk,main="euk")
corrplot(kmer_cor_arch,main="arch")


corrplot(abs(kmer_cor_virus-kmer_cor_arch),is.corr=F,order="hclust")
```

Do we have k-mers highly correlated with the class variable ?

```{r }
kmer_corr_with_class=sort(apply(kmers[,kmer_columns,with=F],MARGIN=2,FUN=function(col){cor(col,as.numeric(factor(kmers$class)))}))
dotchart(kmer_corr_with_class)

```
**TODO** Same on the full sequences, not contigs

Do we have differences in GC?  

```{r }
ggplot(kmers,aes(x=class,y=GC))+geom_boxplot()
ggplot(kmers,aes(x=class,y=sequence_length))+geom_boxplot()+ylim(0,1000)
```


We classify using caret

```{r }
library(caret)
test.ratio=0.8

sub_sample = function(table,size){table[sample(1:nrow(table),size=min(nrow(table),size)),]}

sub_data=rbind(sub_sample(kmers[class=="arch"],1000),sub_sample(kmers[class=="bact"],1000),sub_sample(kmers[class=="virus"],1000),sub_sample(kmers[class=="euk"],1000))

learning_data=data.frame(sub_data[,c("class",kmer_columns),with=F])
learning_data$class=as.numeric(factor(learning_data$class))
inTrain <- createDataPartition(learning_data$class, p = test.ratio, list = FALSE)
trainExpr <- learning_data[inTrain,]
testExpr <- learning_data[-inTrain,]

trainClass <- trainExpr$class
testClass <- testExpr$class

trainExpr$class<-NULL
testExpr$class<-NULL

cat("learning_data and folds generated: ",dim(learning_data),"\n")


# library(doMC)
# registerDoMC(cores = 2)
# registerDoMC(cores = 6)

preProcValues <- preProcess(trainExpr, method = c("center", "scale"))

trainExpr = predict(preProcValues,trainExpr)
testExpr=predict(preProcValues,testExpr)


```

We can launch the training 

```{r}
knn_opts = data.frame(.k=c(9,11,15))
this_train_control = trainControl(method="cv", number=1,verboseIter=T)
print("Start training")
model_performance <- train(x = trainExpr, y = trainClass,method = "knn",trControl=this_train_control,tuneGrid=knn_opts)


print(training_method)
print(model_performance)


final_pred = data.table(extractPrediction(list(model_performance),testX=testExpr,testY=testClass))
knn_cm_train=confusionMatrix(final_pred[dataType=="Training",pred], final_pred[dataType=="Training",obs])
knn_cm_test=confusionMatrix(final_pred[dataType=="Test",pred], final_pred[dataType=="Test",obs])

print(knn_cm_test)
print(knn_cm_train)


```

**TODO** Why fail with caret?

```{r }
library(class)
knn_res=knn(trainExpr,testExpr,trainClass,k=11)
```

**TODO** Very slow !


Using my own code 

```{r }

library(caret)
library(fields)
library(limma)
folds=createFolds(trainClass,k=2)

k_range = c(9,11,13)


## Calculate distance and nearest neighbors
# a_d_mat = 1-cor(t(learning_data))
# # a_d_mat = as.matrix(dist(learning_data))
# # a_d_mat<-KLdiv(t(learning_data))

# rownames(a_d_mat)<-learning_data$domain
# colnames(a_d_mat)<-learning_data$domain
# NN = t(apply(a_d_mat[test_set, training_set], 1, rank))


# Alternative version, where we compute only the distances that we will measure 
# a_d_mat=1-cor(t(learning_data[test_set,]),t(learning_data[training_set,]))


#### More memory savy approach: Columns are sorted and we only keep the top K columns in the NN object 
all_estimates=data.table()
for( i in 1:length(folds)){
	fold=folds[i][[1]]
	this_validation_set= fold
	this_training_set=setdiff(1:length(inTrain), fold)

	a_d_mat=rdist(trainExpr[this_validation_set,],trainExpr[this_training_set,])
	# a_d_mat=1-cor(t(trainExpr[this_validation_set,]),t(trainExpr[this_training_set,]))


	rownames(a_d_mat)<-trainExpr[this_validation_set,]$domain
	colnames(a_d_mat)<-trainExpr[this_training_set,]$domain
	#corrplot(a_d_mat,is.corr=F)

	NN_restricted = as.matrix(t(apply(a_d_mat, 1, order))[,1:max(k_range)],nrows=length(this_validation_set))
	for(k in k_range){
		cat(i,k,"\n")
		these_pred = apply(NN_restricted, 1, function(nn){
		    tab = table(trainClass[this_training_set][nn[1:k]])
		    names(tab)[which.max(tab)]
		})
		est<-confusionMatrix(these_pred,trainClass[this_validation_set])
		est$overall$k=k
		est$overall$fold=i
		all_estimates<-rbind(all_estimates,est$overall)
		# Inspect the results
		# corrplot(table(these_pred, learning_classes[this_validation_set]),is.corr=F,method="square",main=paste("k=",k))
	}
}

print(all_estimates[,mean(Accuracy),by=k])
# We choose the best model
best_k <- all_estimates[,mean(Accuracy),by=k][order(V1,decreasing=T)][1,k]

a_d_mat=rdist(learning_data[-inTrain,2:ncol(learning_data)],learning_data[+inTrain,2:ncol(learning_data)])
# a_d_mat=1-cor(t(learning_data[-inTrain,value_cols]),t(learning_data[+inTrain,value_cols]))


rownames(a_d_mat)<-learning_data[-inTrain,'class']
colnames(a_d_mat)<-learning_data[inTrain,'class']

#corrplot(a_d_mat,is.corr=F)

NN_restricted = as.matrix(t(apply(a_d_mat, 1, order))[,1:best_k],nrows=length(inTrain))
training_labels=learning_data[inTrain]
these_pred = apply(NN_restricted, 1, function(nn){
    tab = table(training_labels[nn[1:k]])
    names(tab)[which.max(tab)]
})
cm_test<-confusionMatrix(these_pred,learning_data[-inTrain,"class"])



```


We compare to a setup with only the most correlated K-mers are choosen

```{r }
# best_kmers=names(head(kmer_corr_with_class,n=3))
best_kmers=names(tail(kmer_corr_with_class,n=3))

a_d_mat=rdist(learning_data[-inTrain,best_kmers],learning_data[+inTrain,best_kmers])
# a_d_mat=1-cor(t(learning_data[-inTrain,value_cols]),t(learning_data[+inTrain,value_cols]))


rownames(a_d_mat)<-learning_data[-inTrain,'class']
colnames(a_d_mat)<-learning_data[inTrain,'class']

#corrplot(a_d_mat,is.corr=F)

NN_restricted = as.matrix(t(apply(a_d_mat, 1, order))[,1:best_k],nrows=length(inTrain))
training_labels=learning_data[inTrain]
these_pred = apply(NN_restricted, 1, function(nn){
    tab = table(training_labels[nn[1:k]])
    names(tab)[which.max(tab)]
})
cm_test<-confusionMatrix(these_pred,learning_data[-inTrain,"class"])
cm_test

```

We see that only selecting a subset of the k-mers (supposedly the most correlated) actually worsens the kappa of the test set 



Trying with recursive trees 

```{r }
library(party)
training_data=cbind(trainExpr,trainClass)
training_data$trainClass <- factor(as.numeric(factor(training_data$trainClass)))

tt=ctree(trainClass~.,data=training_data)
predicted = predict(tt,testExpr)
cm_test<-confusionMatrix(predicted,factor(as.numeric(factor(testClass))))
```