We perform an analysis of virus hardness when we restrict to viruses. 


```{r libraries}
library(reshape2)
library(dplyr)
library(plyr)
library(ggplot2)
library(data.table)
library(stringr)
library(corrplot)
library(caret)

library(foreach)
library(doParallel)
n_cores=2
registerDoParallel(cores=n_cores)

setwd("~/metaviro/experiments/140929 Genebank/")

```
We load the source data 

```{r }
all_instances_kmers=fread("../../data/Genebank/GB_20mb_per_domain_k3.csv")
# all_instances_kmers=fread("../140528_full_ncbi_resampling/full_ncbi_contigs_4k_v2_3mers.csv")
```

* We build a matrix of normalized kmers 

```{r }

kmers_columns=colnames(all_instances_kmers)[5:ncol(all_instances_kmers)]
sequence_attributes=limma::strsplit2(all_instances_kmers$sequence_description,"_",fixed=T)

all_instances_kmers$species=factor(sequence_attributes[,1])
all_instances_kmers$class=factor(sequence_attributes[,2])
all_instances_kmers$contig_index=as.numeric(sequence_attributes[,4])



# all_instances_kmers$bioproject_accession=as.numeric()

all_instances_kmers[,.N,by=class]
```

We restrict to viruses and load the annotations

```{r }
viruses_kmers=all_instances_kmers[class=="virus"] # 10814 contigs 

viruses_bp_accession_numbers=cbind("species"=sequence_attributes[,1],data.table(limma::strsplit2(sequence_attributes[,5],"/|-")))
setnames(viruses_bp_accession_numbers,"V2","BioProject.ID")
viruses_bp_accession_numbers$BioProject.ID=as.integer(as.character(viruses_bp_accession_numbers$BioProject.ID))
viruses_bp_accession_numbers=unique(viruses_bp_accession_numbers[V1=="viruses"])

viruses_kmers=merge(viruses_kmers,viruses_bp_accession_numbers[,list(species,BioProject.ID)],by="species")

virus_annotations=fread("../../data/GENOME_REPORTS/viruses.txt")
setnames(virus_annotations,make.names(colnames(virus_annotations)))
# virus_annotations$BioProject.ID=as.integer(as.character(virus_annotations$BioProject.ID))

annotated_virus_kmers=merge(viruses_kmers,virus_annotations,by="BioProject.ID")

# Any additional duplicates?
annotated_virus_kmers[,.N,by=species][order(N)] # 5539 species, with 1 or 2 contigs each 



```

* Classes of viruses 

```{r }
annotated_virus_kmers[,.N,by='Host'][order(N)] #Vast majority of plants viruses, followed by bacteria then vertebrates. 
annotated_virus_kmers[,.N,by='Group'][order(N)] #Vast majority of plants viruses, followed by bacteria then vertebrates. 
dcast(annotated_virus_kmers,Group~Host)

```

We remove classes with not enough instances 

```{r }
annotated_virus_kmers[Host=="algea",Host:="algae"]
annotated_virus_kmers[Host=="human",Host:="vertebrates, human"]
groups_to_remove=annotated_virus_kmers[,.N,by=list(Group)][order(N)][N<=65,Group]
hosts_to_remove=annotated_virus_kmers[,.N,by="Host"][order(N)][N<=65,Host]
annotated_virus_kmers[Host %in% hosts_to_remove,Host:=NA]
annotated_virus_kmers[Group %in% groups_to_remove,Group:="unclassified"]

dcast(annotated_virus_kmers,Group~Host)
```


We normalize and type 


```{r }
normalized_kmers_counts=annotated_virus_kmers[,kmers_columns,with=F]/annotated_virus_kmers$sequence_length

normalized_kmers=cbind(normalized_kmers_counts, annotated_virus_kmers[,list(species,BioProject.ID,Group,SubGroup,Host,Status,"Segments"=Segmemts,Genes,Proteins,"org"=X.Organism.Name,"org.GC"=GC.,"size"=Size..Kb.)] )

normalized_kmers$org.GC=as.numeric(normalized_kmers$org.GC)
normalized_kmers$size=as.numeric(normalized_kmers$size)
normalized_kmers$Genes=as.numeric(normalized_kmers$Genes)
normalized_kmers$Proteins=as.numeric(normalized_kmers$Proteins)

normalized_kmers$Status=as.factor(normalized_kmers$Status)
normalized_kmers$Group=as.factor(normalized_kmers$Group)
normalized_kmers$SubGroup=as.factor(normalized_kmers$SubGroup)
normalized_kmers$Host=as.factor(normalized_kmers$Host)
normalized_kmers$species=as.factor(normalized_kmers$species)
normalized_kmers$org=as.factor(normalized_kmers$org)
str(normalized_kmers)
```

We plot a PCA of the 3-mers and check any class overlap 

```{r }

pc_kmers=princomp(normalized_kmers_counts)
coord_pca=predict(pc_kmers)[,1:3]
normalized_kmers_pca=cbind(normalized_kmers,coord_pca)

ggplot(normalized_kmers_pca,aes(x=Comp.1,y=Comp.2,colour=Host))+geom_point(alpha=0.5)
```

Any obvious k-mer assoc with the group ?

```{r }
library(party)
t=ctree(Host~org.GC,data=normalized_kmers_pca[!is.na(Host)][1:5000]) # training 
# plot(t)

host_predictions=predict(t,data=normalized_kmers_pca[!is.na(Host)][5001:10000])
confusionMatrix(host_predictions,normalized_kmers_pca[!is.na(Host)][5001:10000,Host])

```

We set up the parallel kNN computation 

```{r }
library(fields)
compute_closest= function(feat_matrix,selected_features,selected_obs,n_closer,selected_labels=NA){
	selected_obs=na.omit(selected_obs)
	dist_m=rdist(feat_matrix[selected_obs,selected_features,with=F],feat_matrix[,selected_features,with=F])
	ranked = apply(dist_m,MARGIN=1,function(r)rank(r,ties.method="min"))
	if(is.na(selected_labels)){
		ldply(1:length(selected_obs),function(i){	
			cat(i,"\n")
			selected_i=which(ranked[,i]<=n_closer); 
			data.table(src=selected_obs[i],tgt=selected_i,rank=ranked[selected_i,i],dist=dist_m[i,selected_i])
		})

	}else{
		ldply(1:length(selected_obs),function(i){	
			selected_i=which(ranked[,i]<=n_closer); 
			tgt_attributes=feat_matrix[selected_i,selected_labels,with=F];
			src_attributes=feat_matrix[selected_obs[i],selected_labels,with=F];
			setnames(src_attributes,paste("src",selected_labels,sep="_"));
			setnames(tgt_attributes,paste("tgt",selected_labels,sep="_"));
			cbind(src_attributes,data.table(src=selected_obs[i],tgt=selected_i,rank=ranked[selected_i,i],dist=dist_m[i,selected_i]),tgt_attributes)
		})
	}
}

```


# Computations 


```{r}

n_closer_neighbors=73
input_size=nrow(normalized_kmers_pca)
nas_to_add=((input_size %/% n_cores)+1)*n_cores - input_size
input_splits= matrix(c(1:input_size,rep(NA,nas_to_add)),byrow=T,ncol=n_cores)

# Single unit test
labels_to_return=c("Group","SubGroup","Host","Status","Segments","Genes","Proteins","org","org.GC","species")
compute_closest(normalized_kmers_pca,kmers_columns,1,50,selected_labels=labels_to_return)

res= data.table(foreach(i=iter(input_splits,by='col'),.combine='rbind',.inorder=F) %dopar% 
	# compute_closest(mixed_3mers,mixed_3mers_count_cols, i[,1],n_closer=n_closer_neighbors,selected_labels=c("class","species","sequence_description")))
	compute_closest(normalized_kmers_pca,kmers_columns,i[,1],n_closer_neighbors,selected_labels=labels_to_return))
save(res,file="73_closes_virus_neighbors.RData")



```






* How many viruses are close to viruses of the same group ? 

```{r }
res_no_diag=res[src!=tgt]
res_no_diag[src==1]

res_neighbors_group=data.table(dcast(res_no_diag,src+src_Group+src_SubGroup~tgt_Group))
res_neighbors_group[,.N,by=src_Group]

res_neighbors_group_m=melt(res_neighbors_group,id.vars=c("src","src_Group","src_SubGroup"))
res_neighbors_group_m[,mismatch:=F]
res_neighbors_group_m[src_Group!=variable,mismatch:=T]

res_neighbors_group_matches=res_neighbors_group_m[,sum(value),by=list(src,src_Group,src_SubGroup,mismatch)]
res_neighbors_group_matches_disagreeing=res_neighbors_group_matches[mismatch==T]
res_neighbors_group_matches_disagreeing[,mean(V1),by=src_Group][order(V1)]

ggplot(res_neighbors_group_matches_disagreeing,aes(x=factor(V1),fill=src_Group))+geom_bar()


```

We restrict to majority classes and perform balanced sampling 

```{r }
majority_group=normalized_kmers_pca[,.N,by=Group][order(N)][N>1000,Group]
# normalized_kmers_pca_maj_group=normalized_kmers_pca[Group %in% majority_group]

# balanced sampling 


normalized_kmers_pca$idx=1:nrow(normalized_kmers_pca)

n_tgt_samples_per_group=normalized_kmers_pca[,.N,by=Group][order(N)][N>1000,min(N)] # 1490 contigs 

downsample_idx=c(
	sample(normalized_kmers_pca[Group == majority_group[1],idx],1490),
	sample(normalized_kmers_pca[Group == majority_group[2],idx],1490),
	sample(normalized_kmers_pca[Group == majority_group[3],idx],1490),
	sample(normalized_kmers_pca[Group == majority_group[4],idx],1490))

normalized_kmers_pca_ds=normalized_kmers_pca[downsample_idx,]
normalized_kmers_pca_ds[,.N,by=Group]
normalized_kmers_pca_ds$idx=1:nrow(normalized_kmers_pca_ds)
```

We compute the closest neighbors restricted to the majority group

```{r }
input_size=nrow(normalized_kmers_pca_ds)
nas_to_add=((input_size %/% n_cores)+1)*n_cores - input_size
input_splits= matrix(c(1:input_size,rep(NA,nas_to_add)),byrow=T,ncol=n_cores)

res_maj_group= data.table(foreach(i=iter(input_splits,by='col'),.combine='rbind',.inorder=F) %dopar% 
	# compute_closest(mixed_3mers,mixed_3mers_count_cols, i[,1],n_closer=n_closer_neighbors,selected_labels=c("class","species","sequence_description")))
	compute_closest(normalized_kmers_pca_ds,kmers_columns,i[,1],n_closer_neighbors,selected_labels=labels_to_return))
save(res_maj_group,file="73_closes_virus_neighbors_maj_groups.RData")
```


We tabulate 

```{r }
res_maj_group_no_diag=res_maj_group[src!=tgt & rank<=53]
# res_maj_group_no_diag[src==1]

res_maj_group_neighbors_group=data.table(dcast(res_maj_group_no_diag,src+src_Group~tgt_Group))
res_maj_group_neighbors_group$src_Group=as.character(res_maj_group_neighbors_group$src_Group)
res_maj_group_neighbors_group[,.N,by=src_Group]


res_maj_group_neighbors_group_m=melt(res_maj_group_neighbors_group,id.vars=c("src","src_Group"))
res_maj_group_neighbors_group_m[,.N,by=src_Group]
res_maj_group_neighbors_group_m[,.N,by=variable]


res_maj_group_neighbors_group_m[,mismatch:=F]
res_maj_group_neighbors_group_m[src_Group!=variable,mismatch:=T]

res_maj_group_neighbors_group_matches=res_maj_group_neighbors_group_m[,sum(value),by=list(src,src_Group,src_SubGroup,mismatch)]
res_maj_group_neighbors_group_matches_disagreeing=res_maj_group_neighbors_group_matches[mismatch==T]
res_maj_group_neighbors_group_matches_disagreeing[,mean(V1),by=src_Group][order(V1)]

group_summaries=res_maj_group_neighbors_group_matches_disagreeing[,sum(V1),by=list(src,src_Group)]


ggplot(group_summaries,aes(x=factor(V1),fill=src_Group))+geom_bar()

```

How many have a majority in their neighbor ? 

```{r }
res_neighbors_group_matches
res_neighbors_group_matches[V1>=36] # more than half 

```


* Are virus subgroup more compact than groups? 

```{r }

most_frequent_subgroups=tail(normalized_kmers_pca[,.N,by=SubGroup][order(N)],n=4)$SubGroup
min_size=min(tail(normalized_kmers_pca[,.N,by=SubGroup][order(N)],n=4)$N)
# Downsample the 6 most frequent subGroups 


downsample_idx=unlist(llply(most_frequent_subgroups,function(g){sample(normalized_kmers_pca[SubGroup == g,idx],min_size)}))
normalized_kmers_pca_ds_SG=normalized_kmers_pca[downsample_idx]
normalized_kmers_pca_ds_SG[,.N,by=Group]
normalized_kmers_pca_ds_SG[,.N,by=SubGroup]


input_size=nrow(normalized_kmers_pca_ds_SG)
nas_to_add=((input_size %/% n_cores)+1)*n_cores - input_size
input_splits= matrix(c(1:input_size,rep(NA,nas_to_add)),byrow=T,ncol=n_cores)

res_maj_sub_group= data.table(foreach(i=iter(input_splits,by='col'),.combine='rbind',.inorder=F) %dopar% 
	compute_closest(normalized_kmers_pca_ds_SG,kmers_columns,i[,1],n_closer_neighbors,selected_labels=labels_to_return))

```


We tabulate 

```{r }
res_maj_sub_group_no_diag=res_maj_sub_group[src!=tgt & rank<=53]
# res_maj_sub_group_no_diag[src==1]

res_maj_sub_group_neighbors_group=data.table(dcast(res_maj_sub_group_no_diag,src+src_Group+src_SubGroup~tgt_SubGroup))
res_maj_sub_group_neighbors_group$src_SubGroup=as.character(res_maj_sub_group_neighbors_group$src_SubGroup)
res_maj_sub_group_neighbors_group[,.N,by=src_SubGroup]


res_maj_sub_group_neighbors_sub_group_m=melt(res_maj_sub_group_neighbors_group,id.vars=c("src","src_Group","src_SubGroup"))
res_maj_sub_group_neighbors_sub_group_m[src==1]
head(res_maj_sub_group_neighbors_sub_group_m[value!=0][order(value)],n=100)
res_maj_sub_group_neighbors_sub_group_m[,.N,by=src_SubGroup]
res_maj_sub_group_neighbors_sub_group_m[,.N,by=variable]
res_maj_sub_group_neighbors_sub_group_m[src %in% c(3,7,8)][order(src)]


res_maj_sub_group_neighbors_sub_group_m[,mismatch:=F]
res_maj_sub_group_neighbors_sub_group_m[src_SubGroup!=variable,mismatch:=T]
head(res_maj_sub_group_neighbors_sub_group_m[mismatch==T][order(src)],n=20)

res_maj_sub_group_neighbors_sub_group_matches=res_maj_sub_group_neighbors_sub_group_m[,sum(value),by=list(src,src_SubGroup,variable,mismatch)]
res_maj_sub_group_neighbors_sub_group_matches_disagreeing=res_maj_sub_group_neighbors_sub_group_matches[mismatch==T]
sub_groups_summaries=res_maj_sub_group_neighbors_sub_group_matches_disagreeing[,sum(V1),by=list(src,src_SubGroup,mismatch)]

ggplot(sub_groups_summaries,aes(x=factor(V1),fill=src_SubGroup))+geom_bar()

```


We PCA and plot 

```{r }
pc_kmers=princomp(normalized_kmers_pca_ds_SG[,kmers_columns,with=F])

coord_pca=predict(pc_kmers)[,1:3]
colnames(coord_pca)=paste("vir_SG",colnames(coord_pca),sep="")

normalized_kmers_pca_ds_SG=cbind(normalized_kmers_pca_ds_SG,coord_pca)

ggplot(normalized_kmers_pca_ds_SG,aes(x=vir_SGComp.1,y=vir_SGComp.2,colour=SubGroup))+geom_point()+facet_wrap(~SubGroup)
ggplot(normalized_kmers_pca_ds_SG,aes(x=vir_SGComp.2,y=vir_SGComp.3,colour=SubGroup))+geom_point()+facet_wrap(~SubGroup)


```


**It seems that sub-groups are very easy to discriminate in euclidean space, we check with a naive kNN** 


```{r }
library(caret)
library(mlbench)

inTrain <- createDataPartition(y = normalized_kmers_pca_ds_SG$SubGroup, p = .9, list = FALSE)
training <- normalized_kmers_pca_ds_SG[ inTrain[,1],c(kmers_columns,"SubGroup"),with=F]
testing  <- normalized_kmers_pca_ds_SG[-inTrain[,1],c(kmers_columns,"SubGroup"),with=F]

knn3_results=knn3Train(training[,kmers_columns,with=F],testing[,kmers_columns,with=F],as.character(training$SubGroup),k=9,prob=F)
confusionMatrix(knn3_results,as.character(testing$SubGroup))


```

Indeed we have pretty good performances. If we instead focus on the four majority group :

```{r }

normalized_kmers_pca$idx=1:nrow(normalized_kmers_pca)
most_frequent_subgroups=tail(normalized_kmers_pca[,.N,by=SubGroup][SubGroup!="unclassified"][order(N)],n=4)$SubGroup
min_size=min(tail(normalized_kmers_pca[,.N,by=SubGroup][SubGroup!="unclassified"][order(N)],n=4)$N)
# Downsample the 6 most frequent subGroups 


downsample_idx=unlist(llply(most_frequent_subgroups,function(g){sample(normalized_kmers_pca[SubGroup == g,idx],min_size)}))
normalized_kmers_pca_ds_SG=normalized_kmers_pca[downsample_idx]
normalized_kmers_pca_ds_SG[,.N,by=Group]
normalized_kmers_pca_ds_SG[,.N,by=SubGroup]

inTrain <- createDataPartition(y = normalized_kmers_pca_ds_SG$SubGroup, p = .8, list = FALSE)
training <- normalized_kmers_pca_ds_SG[ inTrain[,1],c(kmers_columns,"SubGroup"),with=F]
testing  <- normalized_kmers_pca_ds_SG[-inTrain[,1],c(kmers_columns,"SubGroup"),with=F]

knn3_results=knn3Train(training[,kmers_columns,with=F],testing[,kmers_columns,with=F],as.character(training$SubGroup),k=7,prob=F)
confusionMatrix(knn3_results,as.character(testing$SubGroup))


```

We instead try to classify groups 

```{r }
normalized_kmers_pca$idx=1:nrow(normalized_kmers_pca)
most_frequent_groups=tail(normalized_kmers_pca[,.N,by=Group][Group!="unclassified"][order(N)],n=4)$Group
min_size=min(tail(normalized_kmers_pca[,.N,by=Group][Group!="unclassified"][order(N)],n=4)$N)
# Downsample the 6 most frequent Groups 


downsample_idx=unlist(llply(most_frequent_groups,function(g){sample(normalized_kmers_pca[Group == g,idx],min_size)}))
normalized_kmers_pca_ds_SG=normalized_kmers_pca[downsample_idx]
normalized_kmers_pca_ds_SG[,.N,by=Group]

inTrain <- createDataPartition(y = normalized_kmers_pca_ds_SG$Group, p = .8, list = FALSE)
training <- normalized_kmers_pca_ds_SG[ inTrain[,1],c(kmers_columns,"Group"),with=F]
testing  <- normalized_kmers_pca_ds_SG[-inTrain[,1],c(kmers_columns,"Group"),with=F]

knn3_results=knn3Train(training[,kmers_columns,with=F],testing[,kmers_columns,with=F],as.character(training$Group),k=7,prob=F)
confusionMatrix(knn3_results,as.character(testing$Group))

```


Let's add some plant contigs 


```{r }
plant_kmers=all_instances_kmers[class=="euk"]
plant_kmers_ds=plant_kmers[sample(1:nrow(plant_kmers),min_size),]

plant_kmers_ds_counts=plant_kmers_ds[,kmers_columns,with=F]/plant_kmers_ds$sequence_length
plant_kmers_ds_counts$Group="euk"

hybrid_dset=rbind(normalized_kmers_pca[,c(kmers_columns,"Group"),with=F],plant_kmers_ds_counts)
hybrid_dset$idx=1:nrow(hybrid_dset)

# Determine groups to keep
most_frequent_groups=tail(hybrid_dset[,.N,by=Group][Group!="unclassified"][order(N)],n=5)$Group
# Determine minimal number of items per groups 
min_size=min(tail(hybrid_dset[,.N,by=Group][Group!="unclassified"][order(N)],n=5)$N)
# Balanced downsample by groups
downsample_idx=unlist(llply(most_frequent_groups,function(g){sample(hybrid_dset[Group == g,idx],min_size)}))

hybrid_dset_ds=hybrid_dset[downsample_idx]
hybrid_dset_ds[,.N,by=Group]


inTrain <- createDataPartition(y = hybrid_dset_ds$Group, p = .8, list = FALSE)
training <- hybrid_dset_ds[ inTrain[,1],c(kmers_columns,"Group"),with=F]
testing  <- hybrid_dset_ds[-inTrain[,1],c(kmers_columns,"Group"),with=F]

knn3_results=knn3Train(training[,kmers_columns,with=F],testing[,kmers_columns,with=F],as.character(training$Group),k=7,prob=F)
confusionMatrix(knn3_results,as.character(testing$Group))

```





# Determining a p-value of mean class disagreement 

We consider the hardness of the group classification tasks. 

```{r }

most_frequent_groups=tail(normalized_kmers_pca[,.N,by=Group][order(N)],n=4)$Group
min_size=min(tail(normalized_kmers_pca[,.N,by=Group][order(N)],n=4)$N)
# Downsample the 6 most frequent Groups 


downsample_idx=unlist(llply(most_frequent_groups,function(g){sample(normalized_kmers_pca[Group == g,idx],min_size)}))
normalized_kmers_pca_ds_SG=normalized_kmers_pca[downsample_idx]
normalized_kmers_pca_ds_SG[,.N,by=Group]


input_size=nrow(normalized_kmers_pca_ds_SG)
nas_to_add=((input_size %/% n_cores)+1)*n_cores - input_size
input_splits= matrix(c(1:input_size,rep(NA,nas_to_add)),byrow=T,ncol=n_cores)

res_maj_sub_group= data.table(foreach(i=iter(input_splits,by='col'),.combine='rbind',.inorder=F) %dopar% 
	compute_closest(normalized_kmers_pca_ds_SG,kmers_columns,i[,1],n_closer_neighbors,selected_labels=labels_to_return))

```